### 2019-11-12 test.Rmd
Add files via upload

---
title: "Untitled"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)

a <- read.csv("one-beta.csv",sep = ",",skip = 6)
names(a)[names(a) == "mean..count..turtles.on.neighbors..with..ethnicity....ethnicity..of.myself....count..turtles.on.neighbors....of.turtles.with..count..turtles.on.neighbors.....1." ] <- "eth_seg"

a %>% filter(ticks == 0  | ticks == 500  | ticks == 100 | ticks == 1000)  %>%  filter(beta.ie >= 0 & beta.ie <= 15) %>% ggplot(aes(x = beta.ie, y = eth_seg)) + geom_bar(stat="identity") + facet_wrap(~ ticks, labeller = label_both)

b  <-  read.csv("diff_10.csv",sep = ",",skip = 6)
names(b)[names(b) == "mean..count..turtles.on.neighbors..with..ethnicity....ethnicity..of.myself....count..turtles.on.neighbors...of.turtles.with..count..turtles.on.neighbors.....1." ] <- "eth_seg"
names(b)[names(b) == "mean..count..turtles.on.neighbors..with..shape....shape..of.myself....count..turtles.on.neighbors...of.turtles.with..count..turtles.on.neighbors.....1."] <- "val_seg"

c  <- b %>% filter(beta.ie >= beta.iv )

diff_ev <- c$beta.ie - c$beta.iv
c  <- cbind(c,diff_ev)

c %>%  filter(ticks == 1000)  %>% ggplot(aes(x = beta.ie, y = eth_seg)) + geom_bar(stat = "identity") + facet_wrap(~  beta.iv, labeller = label_both)

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.