### 2020-05-02 exp_eth_val.Rmd
Add files via upload

---
output:
  bookdown::pdf_document2:
    toc: false
 #   keep_tex: false
    fig_caption: yes
title: "Observations"
# subtitle: "Target: BIGSSS Research Day"
author: Rocco Paolillo

# date: JU 
#language: en-GB
#bibliography: "references.bib"
# always_allow_html: yes

# target: BIGSSS Research Day by September 15th, presented  on October 25th. 6000 words: around 12 pages.
# csl: apa.csl # by default: Chicago style
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
- \usepackage{amsmath}
- \usepackage{graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.lp="fig:",warning = FALSE)

library(knitr)
library(ggplot2)
library(forcats)
library(reshape2)
library(tidyr)
library(dplyr)
library(reshape)
library(ggpubr)
library(magick)
library(gridExtra)
library(grid)

#theme_set(theme_pubr())

wd <- setwd("C:/Users/rocpa/OneDrive/Desktop/AF/")
dataone <- read.csv("experiments/80_50_onebeta.csv",sep = ";",skip = 1)
datatwo <- read.csv("experiments/two_beta.csv", sep = ";",skip = 1)

dataone$rev_circle_orange <- fct_rev(factor(dataone$circle_orange))
dataone$rev_circle_blue <- fct_rev(factor(dataone$circle_blue))

datatwo$rev_beta_iv_circle <- fct_rev(factor(datatwo$beta_iv_circle))
datatwo$rev_beta_ie_square <- fct_rev(factor(datatwo$beta_ie_square))

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")
 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")
 circle_orange.labs <-  c("20% Circle \nOrange", "50% Circle \nOrange","80% Circle \nOrange")
names(circle_orange.labs) <- c("20", "50","80")

```

# Recap

- Research question: How the interplay of ethnic and value homophily preferences can explain difference in the degree of integration for the members of the same ethnic group (hybrid segregation) for both ethnic and value segregation.
- Focus: different weight on the two dimensions, heterogeneity among agents, relative group size for both dimensions \par

Utility of agents in Schelling relocation

$$
U= \beta_eX_v + \beta_vX_v + \epsilon
$$

```{r thheory, fig.pos="h", out.width="50%",out.width='\\textwidth',fig.align='left',fig.show="hold", fig.cap="Agents distinction and utility function", echo=FALSE, include=TRUE }

knitr::include_graphics("recapp.png")
```

# First Section (effect of relative group size)

* Aim:
    + Effect of different homophily preferences: independent effect of $\beta_e$ on ethnic segregation, and direct effect of beta value $\beta_v$ on value segregation.

Table \ref{tab:first}

\begin{table}[H]
\begin{tabular}{llll}
\midrule
 Static agent variable                      & Range         &      &                \\ \midrule
 Ethnicity  (\textit{color})         & Blue, Orange        \\
 Value orientation (\textit{shape}) & Square, Circle &  &  \\
 \midrule

 Parameters kept constant    & Slider & Range   & Experiments  \\ 
 \midrule
 Population density    & density                 & $[0,0.99]$   & $0.7$          \\
 Peak ethnic utility & similar\_wanted\_ethnic & $[0,1]$ & $1$ (linear function) \\
 Peak value utility & similar\_wanted\_value & $[0,1]$ & $1$ (linear function) \\
 \midrule
 Parameters modified in experiment & Slider  & Range   & Experiments  \\ 
 \midrule
 Ethnic group ratio (local\/minority)  & fraction\_blue &  $[0,1]$      & $0.5,0.8$  \\
 value ratio  blue population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 value ratio  orange population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 Determinism ethnic utility ($\beta_e$) & beta-ie for all agents & $[0,1]$ & $[0,10]$  \\
 Determinism value utility ($\beta_v$) & beta-iv for all agents & $[0,1]$ & $[0,10]$ \\
 \midrule
 Global output measures    & Computation &   Range & \\ 
 \midrule
 Ethnic segregation  & same color neighborhood proportion & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
 Value segregation  & same shape neighborhood proportion & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
   \midrule
\end{tabular}
\label{tab:first}
 \caption{Basic Model}

\end{table}

```{r global, fig.height=8.8, fig.pos="h", include=TRUE, fig.cap="Global Segregation with ethnic equal group size, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

circle_local.labs <- c("20% Local Circle", "50% Local Circle","80% Local Circle")
 names(circle_local.labs) <- c("20", "50","80")
circle_minority.labs <-  c("20% Minority Circle", "50% Minority Circle","80% Minority Circle")
 names(circle_minority.labs) <- c("20", "50","80")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

eth_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs ))  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))   + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
   geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

val_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs )) + 
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

glb <- annotate_figure(ggarrange(eth_eq , ggarrange(val_eq), nrow = 2), top = text_grob("Global Segregation, Majority:Minority = 0.5:0.5",face = "bold", size = 12))

glb

```

```{r global-maj, fig.height=8.8, fig.pos="h",  include=TRUE, fig.cap="Global Segregation with ethnic equal group size, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

 circle_local.labs <- c("20% Local Circle", "50% Local Circle","80% Local Circle")
names(circle_local.labs) <- c("20", "50","80")
 circle_minority.labs <-  c("20% Minority Circle", "50% Minority Circle","80% Minority Circle")
names(circle_minority.labs) <- c("20", "50","80")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

maj_eth <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs ))  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10)) + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_val <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs )) + 
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
 scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_glb <- annotate_figure(ggarrange(maj_eth , ggarrange(maj_val), nrow = 2), top = text_grob("Global Segregation, Majority:Minority = 0.8:0.2",face = "bold", size = 12))
maj_glb

```

```{r baselineone,fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap=" Baseline space", echo=FALSE, include=TRUE  }

knitr::include_graphics("baseline_1.png")

```

```{r crlo,fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, first section", echo=FALSE, include=TRUE  }

crloeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

crlomj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

 crloeq
 crlomj
```

```{r crlblmajority, fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="Circle in majority condition.", echo=FALSE, include=TRUE}

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

```

# Effect of value distribution in group value differences

Each graph is one level of $\%$ circle (value 1) in the local population (majority = blue agents). Each condition panel refers to distribution of  circle agents in the minority population.

```{r heatmapdsquare,fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Square orange, heterogenous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle \nß value 10" ,"Circle \nß value 5", "Circle \nß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_sq_or", "VALUE" = "val_sq_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_iv_square, y = beta_ie_square, fill = Segra)) + geom_tile() + facet_grid(rev_beta_iv_circle ~  beta_ie_circle + segregation, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs, beta_ie_circle = beta_ie_circle.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß value square") + ylab("ß ethnic square") + ggtitle("Square orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

```{r heatmapdcircle, fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, heterogeneous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5","Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")
beta_iv_square.labs <- c("Square \nß value 0", "Square \nß value 5","Square \nß value 10")
names(beta_iv_square.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_cl_or", "VALUE" = "val_cl_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_ie_circle, y = beta_iv_circle, fill = Segra)) + geom_tile() + facet_grid(rev_beta_ie_square ~  beta_iv_square + segregation, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = beta_iv_square.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß ethnic circle") + ylab("ß value circle") + ggtitle("Circle orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

```{r betatwoethnic,fig.pos='t',out.width="96%",fig.show = "hold",fig.align='center',fig.cap="Difference segregation", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle ß value 10" ,"Circle ß value 5", "Circle ß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

#  v_circle.labs <- c("ß value circle 0", "ß value circle 5", "ß value circle 10")
# names(v_circle.labs) <- c("0", "5","10")
e_circle.labs <- c("Circle ß ethnic 0", "Circle ß ethnic 5", "Circle ß ethnic 10")
names(e_circle.labs) <- c("0", "5","10")

ie_10 <- circle_frame %>%  filter(beta_ie_square == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point() + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß ethnic square 10 - ß value square", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5 <- circle_frame %>%  filter(beta_ie_square == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point() + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  scale_x_discrete(name = "ß ethnic square 5 - ß value square", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5
ie_10
```

```{r betatwovalue,  fig.pos="t",fig.show = "hold", fig.align='center',fig.show="hold", fig.cap="Difference segregation between", echo=FALSE, include=TRUE  }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5", "Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")

e_square.labs <- c("ß ethnic \nsquare 0", "ß ethnic \nsquare 5", "ß ethnic \nsquare 10")
names(e_square.labs) <- c("0", "5","10")
v_square.labs <- c("ß value square 0", "ß value square 5", "ß value square 10")
names(v_square.labs) <- c("0", "5","10")

iv_10 <- circle_frame %>%  filter(beta_iv_circle == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point() + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5 <- circle_frame %>%  filter(beta_iv_circle == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point() + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5
iv_10

```

# ANNEX

```{r sqro, fig.pos='t',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Orange", echo=FALSE, include=TRUE  }

sqroeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroqmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroeq
sqroqmj
```

```{r crlbl,fig.pos='t' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Circle Blue", echo=FALSE, include=TRUE}

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")

crlbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

crlbmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

 crlbeq
 crlbmj
```

```{r sqrbl, fig.pos='t' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Blue", echo=FALSE, include=TRUE}

sqrbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbomj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbeq
sqrbomj
```