### 2020-05-06 exp_eth_val.Rmd
Add files via upload

---
output:
  bookdown::pdf_document2:
    toc: false
 #   keep_tex: false
    fig_caption: yes
title: "Observations"
# subtitle: "Target: BIGSSS Research Day"
author: Rocco Paolillo

# date: JU 
#language: en-GB
bibliography: "references.bib"
# always_allow_html: yes

# target: BIGSSS Research Day by September 15th, presented  on October 25th. 6000 words: around 12 pages.
# csl: apa.csl # by default: Chicago style
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
- \usepackage{amsmath}
- \usepackage{graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.lp="fig:"  ,warning = FALSE)

library(knitr)
library(ggplot2)
library(forcats)
library(reshape2)
library(tidyr)
library(dplyr)
library(reshape)
library(ggpubr)
library(magick)
library(gridExtra)
library(grid)

#theme_set(theme_pubr())

wd <- setwd("C:/Users/rocpa/OneDrive/Desktop/AF/")
dataone <- read.csv("experiments/80_50_onebeta.csv",sep = ";",skip = 1)
datatwo <- read.csv("experiments/two_beta.csv", sep = ";",skip = 1)

dataone$rev_circle_orange <- fct_rev(factor(dataone$circle_orange))
dataone$rev_circle_blue <- fct_rev(factor(dataone$circle_blue))

datatwo$rev_beta_iv_circle <- fct_rev(factor(datatwo$beta_iv_circle))
datatwo$rev_beta_ie_square <- fct_rev(factor(datatwo$beta_ie_square))

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")
 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")
 circle_orange.labs <-  c("20% Circle \nOrange", "50% Circle \nOrange","80% Circle \nOrange")
names(circle_orange.labs) <- c("20", "50","80")

```
\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}

# Recap

- Research question: How the interplay of ethnic and value homophily preferences can explain difference in the degree of integration for the members of the same ethnic group (hybrid segregation) for both ethnic and value segregation.
- Focus: different weight on the two dimensions, heterogeneity among agents, relative group size for both dimensions \par

Utility of agents in Schelling relocation

$$
U= \beta_eX_v + \beta_vX_v + \epsilon
$$

```{r thheory, fig.pos="h", out.width="50%",out.width='\\textwidth',fig.align='left',fig.show="hold", fig.cap="Agents distinction and utility function", echo=FALSE, include=TRUE }

knitr::include_graphics("recapp.png")
```

# First Section (effect of relative group size)

* Aim:
    + Effect of different homophily preferences: direct effect of $\beta_e$ on ethnic segregation, and direct effect of beta value $\beta_v$ on value segregation, and reciprocal influences, i.e. how of $\beta_v$ influences effect of  $\beta_e$ on ethnic segregation and how $\beta_e$ influences effect of $\beta_v$ on value segregation
* Focus: interaction between relative group size for each dimension with degree of $\beta$, since group size directly influence the probability to find similar ones and the systematic utility computed for each neighborhood
* Assumption: 
    + not heterogeneity: all agents hold both ethnic preference (same color) and value preference (same shape)
    + ethnic size at global level: equal size (blue/orange: $0.50$), majority/minority (blue/orange: $0.80$)
    + value size within each ethnic group: minority circle circle/square $0.20$, equal size circle/square $0.50$, majority circle circle/square $0.80$, imagining tolerant (circle) and intolerant (square),this means to move from more tolerant to less tolerant groups, both majority and minority. In this section there is not difference in preferences implied Tab \ref{tab:a}
* Figures:
    + Fig: \ref{fig:baselineone} represents how results are presented through the space of the model. Axis represent the ratio of $circle/square$ in each ethnic group, expressed as percentage: on y-axis the percentage of circle value in the orange ethnic group, on the x-axis the percentage of circle value in the blue ethnic group. The percentage of square agents can be calculated from it. The \textbf{baseline} condition is represented by the black dot, where each ethnicity is split equally into the two value groups ($0.50$ circle, $0.50$ square); following the diagonal, each ethnic group moves from minority circle to majority circle. Each obervation is repeated  for ethnic equal group sizes (blue/orange = $0.50$) and ethnic majority/minority condition (blue/orange = $0.80$). In the baseline condition for equal ethnic size, each group $valueXethnicity$ represents the $25 \%$ of the overall population. This means each agent in Fig: \ref{fig:thheory} has potentially probability $0.25$ to relocate close to someone who shares both ethnicity and value orientation.
    + In all results $\beta = 5$ seems a threshold between area of high randomness and low randomness. I would use this as reference point in describing figures and change in conditions

\begin{table}[H]
\begin{tabular}{llll}
\midrule
 Static agent variable                      & Range         &      &                \\ \midrule
 Ethnicity  (\textit{color})         & Blue, Orange        \\
 Value orientation (\textit{shape}) & Square, Circle &  &  \\
 \midrule

 Parameters kept constant    & Slider & Range   & Experiments  \\ 
 \midrule
 Population density    & density                 & $[0,0.99]$   & $0.7$          \\
 Peak ethnic utility & similar\_wanted\_ethnic & $[0,1]$ & $1$ (linear function) \\
 Peak value utility & similar\_wanted\_value & $[0,1]$ & $1$ (linear function) \\
 \midrule
 \textbf{Parameters modified in experiment} & \textbf{Slider} & \textbf{Range}   & \textbf{Experiments}  \\ 
 \midrule
 Ethnic group ratio (local\/minority)  & fraction\_blue &  $[0,1]$      & $0.5,0.8$  \\
 value ratio  blue population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 value ratio  orange population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 Determinism ethnic utility ($\beta_e$) & beta-ie for all agents & $[0,10]$ & $[0,10]$  \\
 Determinism value utility ($\beta_v$) & beta-iv for all agents & $[0,10]$ & $[0,10]$ \\
 \midrule
 Global output measures    & Computation &   Range & \\ 
 \midrule
 Ethnic segregation  & same color neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
 Value segregation  & same shape neighborhood $\%$  & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
   \midrule
Number of Runs   & 1 & Numer of steps & 1000 \\ 
 \midrule
\end{tabular}
 \caption{First Section} 
 \label{tab:First}
\end{table}

## Global Observations
* Fig: \ref{fig:global} and  Fig: \ref{fig:global-maj} shows the global ethnic and value segregation, for equal ethnic group sizes in Fig: \ref{fig:global}  and majority/minority ethnic condition in Fig: \ref{fig:global-maj}  . The heatmaps report in each condition  $\beta_e$ on x-axis, and  $\beta_v$ on y-axis. 
* Ethnic segregation increases along the axis of $\beta_e$, while value segregation symmetrically along the axis of $\beta_v$. In both cases $\beta = 5$ shows a threshold between areas of integration and area of segregation.
* While area below $\beta = 5$ is linearly increasing, areas above $\beta = 5$ seem to show same degree of high segregation. Attributing this to randomness component, I would identify area $\beta < 5$ as low randomness, area  $\beta \geq 5$ high randomness, and focus on threshold of $\beta = 5$, this helps presenting figures later
* The threshold of $\beta = 5$ to me can relate to "sensitivity to change" in @van2009neighborhood: since the shape of utility is linearly increasing, each change in the composition of neighborhood will increase the difference between options. Once the difference is multiplied times $\beta$, the effect of utility difference is too high and soon the best option is taken. This means that only for very low levels of $\beta$ it is possible to see a linear increasing in the systematic utility. At $\beta = 5$ the systematic utility becomes too strong and agents highly sensitive to the change in neighborhood composition, moving to the best solution, increasing segregation of neighborhoods. This soon attires other agents and equilibria towards segregation are easily reached.
* The direct effect of $\beta_e$ on ethnic segregation and  $\beta_v$ on value segregation seems unaffected by the other systemic component, both when  $\beta = 0$, so that agents relocate randomly for value or ethnic homophily, and when they increase the systemic utility for either ethnic (increase in  $\beta_e$) or value homophily (increase in $\beta_v$)
* Condition of ethnic majority/minority shows basically the same patterns, slighter high ethnic segregation for the area of high randomness, which is direct effect of unequal group sizes.
* Change in ethnic segregation compared to the baseline seem to occur only along the diagonal, where circle orientation represents the minority in one group and majority in the other, or vice versa. Similar slightly changes occur in the area of high randomness $\beta_e < 5$ with increase of $\beta_v$, which is counterintuitive. Circle of both ethnic group not care about ethnic composition of the neighborhood, but they care about the value composition. Since in one ethnic group circle represent the value majority, and in the other the circle minority, circle in the minority condition will result in ethnically mixed or majority out-group neighborhoods, circle in the majority in ethnically homogeneous neighborhoods, since they have more probability to find someone circle of the own ethnic group than the other. So, for these agents ethnic segregation is the more likely outcome, also if the systematic component of utility is for value segregation and ethnic utility random. The global results are influenced by the numeric superiority of these  agents.
    * Next figures look at the behavior of specific subgroup in Fig. \ref{fig:thheory}

```{r baselineone,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap=" Baseline space", echo=FALSE, include=TRUE  }

knitr::include_graphics("baseline_1.png")

```

```{r global, fig.height=8.8, fig.pos="H", include=TRUE, fig.cap="Global Segregation with ethnic equal group size, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

 circle_local.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_local.labs) <- c("20", "50","80")
 circle_minority.labs <-  c("20% Circle \nOrange","50% Circle \nOrange","80% Circle \nOrange")
names(circle_minority.labs) <- c("20", "50","80")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

eth_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs ))  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))   + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
   geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

val_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs )) + 
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

glb <- annotate_figure(ggarrange(eth_eq , ggarrange(val_eq), nrow = 2), top = text_grob("Global Segregation, Blue:Orange = 0.5:0.5",face = "bold", size = 12))

glb

```

```{r global-maj, fig.height=8.8, fig.pos="H",  include=TRUE, fig.cap="Global Segregation with majority/minority ethnicity, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

 circle_local.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_local.labs) <- c("20", "50","80")
 circle_minority.labs <-  c("20% Circle \nOrange","50% Circle \nOrange","80% Circle \nOrange")
names(circle_minority.labs) <- c("20", "50","80")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

maj_eth <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs ))  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10)) + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_val <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  circle_orange = circle_minority.labs )) + 
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
 scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10))  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_glb <- annotate_figure(ggarrange(maj_eth , ggarrange(maj_val), nrow = 2), top = text_grob("Global Segregation, Blue:Orange = 0.8:0.2",face = "bold", size = 12))
maj_glb

```

## Observations for group $ethnicityXvalue$:

* Figures \ref{fig:crlo}, \ref{fig:crlominority} and \ref{fig:crlblmajority} replicate the behavior space in Fig: \ref{fig:baselineone} used in the previous section. For each condition the ethnic and value segregation are reported.
* I report the results of circle orange in the condition of ethnic equal size, ethnic minority and ethnic majority. Since agents  follow the same homophily preferences, the only difference is due to the manipulation of relative group sizes for ethnicity and value orientation, so that the results here presented can be applied to all other group by rotating the axis on Fig: \ref{fig:baselineone} or results along the diagonal. I do this in the ANNEX as demonstration
* 3 outcomes are evident, rather than 2 as in traditional assessment of Schelling: integration (filler white  0.5), segregation (filler red above 0.5), and assimilation (filler blue close to 0, only members outgroup). This can relate as measure of *hybrid segregation*

### Equal ethnic group size (Fig: \ref{fig:crlo})

* Given ethnic equal group size, at the baseline condition ($50 \%$ circle orange, $50 \%$ circle blue), each group replicates what observed for global segregation: high randomness close to integration for $\beta < 5$, low randomness high segregation for $\beta \geq 5$. For ethnic segregation this is the same along the diagonal, given equal ethnic sizes
* On the top left corner,  where the value group represents majority in the own ethnic group, but minority in the other, integration is the most likely outcome  only for low levels of both $\beta_e$ and  $\beta_v$ (low left corner), with middle segregation in the rest of high randomness area. While the value becomes more represented in the other ethnic group, ethnic integration increases  with increase of  $\beta_v$ in the area of high randomness ($\beta_e < 5$), untouched in the area of low  randomess ($\beta_e \geq 5$). In the same conditions, high value segregation appears as direct effect of increase in the proportion of same value oriented circle in the other group
* Low row, where the value group represents minority in the own group: value assimilation to square is the case when circle is the minority in both etnic groups. Still, comparing to ethnic segregation, area of high randomenss appears ethnically integrated, area of low randomness ethnically segregated. This implies that for  low $\beta_e$ and low $\beta_v$ value minority relocate in ethnically mixed neighborhoods, but where square is the majority. For high  $\beta_e$, circle live  in ethnically segregated neighborhoods, where still value represent the majority of the own ethnic group. Though low $\beta_e$ reproduces the  initial random ethnic distribution, this doesn't happen for value distribution due to effect of majority size of square in both groups.
* Proceeding with increase of same  value in the other ethnic group (x-axis), value assimilation increases for high levels of $\beta_e$ and low levels of $\beta_v$, in the same conditions ethnic segregation. As before circle agents increase  ethnic utility, but still have more chance to relocate close to square co-ethnics. Along $\beta_v$, increase implies value segregation as expected,  but ethnic assimilation when $\beta_e$ is low: circle agents not care about ethnic utility, but maxize value orientation, so that most appealing neighborhoods are those inhabited by circle of the other group.
* The last observation is noteworthy to me, though predictable: as alternative to assimilation due to being in the ethnic minority size, people can fall into ethnic (spatial) assimilation because of an other characteristic (value) which is unequally represented in two ethnic group numerically equal

```{r crlo,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, equal ethnic size", echo=FALSE, include=TRUE  }

dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

```

### Ethnic minority (Fig: \ref{fig:crlominority})

* Trends are generally similar to Fig: \ref{fig:crlo}
* In the baseline, ethnic assimilation emerges instead of ethnic integration, becuase of the ethnic minority of the group.
* Up left corner, where circle represent the majority value of ethnic  minority and the minority value of ethnic majority, compared to Fig: \ref{fig:crlo} ethnic assimilation rather than integration emerges for the area of high randomness for both dimensions, and ethnic integration rather than segregation for the increase of $\beta_v$ and low values of $\beta_e$. While for areas of high randomness agents fall into assimilation because of numeric inferiority, when they increase value utility $\beta_v$ they want to relocate close to other circle agents, regardless of their ethnicity. Since in this condition circle represent the minority value of the majority ethnic population (the opposite of circle orange in ethnic minority), the combination allows for the increase in value preference to make up for the ethnic numeric inferiority, so that both circle minority and circle majority end up in neighborhoods ethnically integrated and value circle homogeneous (cfr Fig: \ref{fig:crlbl}, right down corner).
* Low row: the effect of increase in $\beta_e$ on value assimilation is similar as Fig: \ref{fig:crlo}, since circle here represents the minority in the own group. Increase of $\beta_v$ shows instead higher assimilation, because circle minority has more chance to relocate close to similar circle value of the other ethnic group, proportionally to value circle being the majority value in the majority ethnic group. Interestingly, given the area of high ethnic utility for $\beta_e \geq 5$, at the increase of $\beta_v$ less homogenous ethnic segregation emerges compared to the equal ethnic size.
* The last two  points are noteworthy to me: the assumption that similar values can foster ethnic integration over ethnic segregation can work provided that an adjustement between distribution of values in both ethnic group and preferences for ethnic or value utility are  taken into account.

```{r crlominority,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, ethnic minority", echo=FALSE, include=TRUE  }

dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

```

### Ethnic majority (Fig: \ref{fig:crlblmajority})
* Low row where circle are the minority in the ethnic majority blue, show the same pattern for the circle in minority and equal ethnic size: increasing in ethnic utility implies value assimilation to square  members of the same ethnic group, because of value numeric minority
* Value integration occurs in the upper row ($50 \% $ circle in blue majority ethnic) for random area of low $\beta_v$, because eqaul chance to relocate close to either circle or square, but in both cases of same ethnic group
* Compared to the baseline in Fig: \ref{fig:crlo}, ethnic segregation emerges also in area of high randomness ($\beta_e < 5$), because of numeric superiority, also for the upper side of the diagonal
* Low right corner, where circle are the minority in the blue ethnicity, but majority in the ethnic minority, ethnic integration emerges for high values of $\beta_v$, despite ethnic minority of orange. Same as in Fig: \ref{fig:crlominority}

```{r crlblmajority, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="Circle blue, ethnic majority", echo=FALSE, include=TRUE}

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

```

# Second Section (heterogeneous preferences for tolerant and intolerants)

* Aim:
    + Test the effect of heterogeneous preferences for two value groups, i.e. circle value-oriented care more about value than ethnic neighborhood composition, square ethnicity-oriented care more about ethnic than value composition
    + This would relate to the difference between tolerant (circle) and intolerant (square) as in the ACS paper and that we discussed
    + Observe how heterogeneity in agents' preferences along two value oriented groups will affect results in the previous section, through their mutual adaptation
    + Interaction between heterogeneous preferences and relative group sizes for both dimensions (here only equal ethnic and value group)
* Assumption: 
    + Agents hold both ethnic and value preferences, but $\beta$ for each dimension depends on their value orientation, either circle or square. Not difference due to ethnic membership are taken into consideration
    + Given the linear function, utility increases with value or ethnic homogeneity of neighborhood, but determinism component $\beta_e$ and $\beta_v$ are differently distributed to value groups: ethnic preferences of circle cannot be higher than value preferences ($\beta^{\circ}_e \leq \beta^{\circ}_v$) and value preferences of square cannot be higher than thier value preferences ($\beta^{\square}_v \leq \beta^{\square}_e$)
    + Ideally also condition of ethnic majority/minority and circle majority/minority should be included. Here for matter of time it is reduced to equal ethnic group size (blue/orange $0.50$) and for each ethnic group equal value group (circle/square $0.50$). So, each of the four subgroup in Fig: \ref{fig:thheory} represents the $25 \%$ of the global agents' population
* Figures:
    + I compare the ethnic and value segregation of both circle and square agents
    + Focus on the difference between ethnic and value segregation (being close to agents sharing both ethnicity and value)
    + To reduce observations given the numbers of parameters, I jump on 3 critical points individualized in the previous section: $\beta = 0$: total randomness, $\beta = 5$: min. determinism, $\beta = 10$: max determinism
    + For each value group, in the baseline I put on x-axis the difference between the highest reference $\beta$ of that group and the other one, i.e. $\beta_e - \beta_v$ for square, $\beta_v - \beta_e$ for circle. Considering the critical point, I repeat observation for reference $\beta = 5$: min. determinism, and $\beta = 10$: max determinism, $\beta = 0$ is included in bot. Thus, along the x-axis the group becomes more sensitive to the reference dimension, ethnicity for square group, value for circle group. I compare how the baseline condition changes for interaction of value and ethnic preferences of the other group faceting the 3 critical points $\beta = 0; \beta = 5;\beta = 10$

\begin{table}[H]
\begin{tabular}{llll}
\midrule
 Static agent variable                      & Range         &      &                \\ \midrule
 Ethnicity  (\textit{color})         & Blue, Orange        \\
 Value orientation (\textit{shape}) & Square, Circle &  &  \\
 \midrule

 Parameters kept constant    & Slider & Range   & Experiments  \\ 
 \midrule
 Population density    & density                 & $[0,0.99]$   & $0.7$          \\
 Peak ethnic utility & similar\_wanted\_ethnic & $[0,1]$ & $1$ (linear function) \\
 Peak value utility & similar\_wanted\_value & $[0,1]$ & $1$ (linear function) \\
 \midrule
 \textbf{Parameters modified in experiment} & \textbf{Slider} & \textbf{Range}   & \textbf{Experiments}  \\ 
 \midrule
 Ethnic group ratio (local\/minority)  & fraction\_blue &  $[0,1]$      & $0.5$  \\
 value ratio  blue population & circle\_blue & $[0,1]$      & $0.5$  \\
 value ratio  orange population & circle\_blue & $[0,1]$      & $0.5$  \\
 Determinism ethnic utility square ($\beta^{\square}_e$) & beta-ie-sqr & $[0,10]$ & $[0,10]$  \\
 Determinism value utility square ($\beta^{\square}_v$) & beta-iv-sqr  & $[0,10]$ & $[0,10], if \beta^{\square}_v \leq \beta^{\square}_e$ \\
 Determinism value utility circle ($\beta^{\circ}_v$) & beta-iv-crl  & $[0,10]$ & $[0,10]$  \\
 Determinism ethnic utility circle ($\beta^{\circ}_e$) & beta-ie-crl & $[0,10]$ & $[0,10], if \beta^{\circ}_e \leq \beta^{\circ}_v$  \\
 \midrule
 Global output measures    & Computation &   Range & \\ 
 \midrule
 Ethnic segregation  & same color neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
 Value segregation  & same shape neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
   \midrule
Number of Runs   & 1 & Number of steps & 1000 \\ 
 \midrule
\end{tabular}
 \caption{Second section: heterogeneous preferences} 
 \label{tab:Second}
\end{table}

## Comparison different value orientation
* Fig: \ref{fig:betatwoethnic}: x-axis $\beta^{\square}_e - \beta^{\square}_v$, on the top panel with $\beta^{\square}_e = 5$ (area of high randomness), on bottom panel  $\beta^{\square}_e = 10$ (area of low randomness). X-axis = 0,  ethnic and value utility have equal weight, along the x-axis ethnic weight prevails over value weight. the lower the distance between ethnic and value segregation, the more agents live close to others similar for both ethnicity and value.
* Viewing at lower left panel, where behavior of circle is totally random, not only the distance between ethnic and value segregation of square is lower when $\beta^{\square}_e = \beta^{\square}_v$ (x-axis 0), but also higher for both dimensions, progressing towards x-axis (ethnic becomes higher than 10), value segregation decreases, with more linear slope for high determinism area (bottom panel).
* It persists that each $\beta$ rules the reference dimension, but the random behavior of the other group decreases the segregation expected for that level of $\beta$, higher determinism of the other group increases  the level of segregation for that dimension. Demonstration, moving from $\beta^{\circ}_v = 5$ to  $\beta^{\circ}_v = 10$ 
* As $\beta^{\square}_e$ becomes higher than  $\beta^{\square}_v$, moving on the x-axis, ethnic segregation persists constantly, while value segregation diminishes.
* The threshold of $\beta = 5$ seen in the previous section is somehow replicated: in the top panel, max value and ethnic segregation are equal 1 when both  $\beta^{\square}_e = \beta^{\square}_v = 0$, on the x-axis 0, and value segregation declines once $\beta^{\square}_v$ comes close to 0. Looking at panel below (high determinism), the decline of value segregation occurs equally when $\beta^{\square}_v$ declines from 5 to 0 (on the right of red line 10-5:5 and so on)
* Even though the direct effect of $\beta_v$ is assumed and same patterns are observed from low determinism (top panel) and high determinism (bottom panel), the high determinism of  $\beta^{\square}_e$ is associated with very slight increase in the value segregation of square, also for same level of $\beta^{\square}_v = 0$, though negligeable
* The distance between ethnic and value segregation is higher in the middle panel, where circle hold same min utility preferences $\beta^{\circ}_e = \beta^{\circ}_v = 5$, and also square increase in ethnic over value preference (x-axis = 5). Basically the only random component is the value orientation of square agents, which has the same effect on both circle and square population patterns
* Moving on the diagonal, where $\beta^{\circ}_e = \beta^{\circ}_v = 10$, the distance between ethnic and value decreases, meaning agents are more likely to live close to agents of both same ethnicity and value: the increase in utility at max level of circle ($\beta^{\circ}_e = \beta^{\circ}_v = 10$) compensates for random value behavior of square, since circle will avoid neighborhoods with square 
* Fig: \ref{fig:betatwovalue} shows the similar trends, focusing on value segregation over ethnic segregation for the circle population. Nevertheless, in the middle panel ($\beta^{\square}_e = \beta^{\square}_v = 5$), I would expect symmetric scenario than in Fig: \ref{fig:betatwoethnic}, with ethnic segregation of circle being same as value segregatiion than square. Why here circle are completed ethnically integrared? There is a difference between the two dimensions? Think about.

```{r betatwoethnic,fig.pos='H',fig.show = "hold",fig.align='center',fig.cap="Difference segregation, x-axis: difference ethnic and value preference square, faceting conditions: preferences of circle", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle ß value 10" ,"Circle ß value 5", "Circle ß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

#  v_circle.labs <- c("ß value circle 0", "ß value circle 5", "ß value circle 10")
# names(v_circle.labs) <- c("0", "5","10")
e_circle.labs <- c("Circle ß ethnic 0", "Circle ß ethnic 5", "Circle ß ethnic 10")
names(e_circle.labs) <- c("0", "5","10")

ie_10 <- circle_frame %>%  filter(beta_ie_square == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß ethnic square 10 - ß value square", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5 <- circle_frame %>%  filter(beta_ie_square == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  scale_x_discrete(name = "ß ethnic square 5 - ß value square", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5
ie_10
```

```{r betatwovalue,  fig.pos="H",fig.show = "hold", fig.align='center',fig.show="hold", fig.cap="Difference segregation between", echo=FALSE, include=TRUE  }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5", "Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")

e_square.labs <- c("ß ethnic \nsquare 0", "ß ethnic \nsquare 5", "ß ethnic \nsquare 10")
names(e_square.labs) <- c("0", "5","10")
v_square.labs <- c("ß value square 0", "ß value square 5", "ß value square 10")
names(v_square.labs) <- c("0", "5","10")

iv_10 <- circle_frame %>%  filter(beta_iv_circle == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5 <- circle_frame %>%  filter(beta_iv_circle == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5
iv_10

```

# ANNEX A {-}

```{r sqro, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Orange", echo=FALSE, include=TRUE  }

sqroeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroqmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroeq
sqroqmj
```

```{r crlbl,fig.pos='H' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Circle Blue", echo=FALSE, include=TRUE}

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")

crlbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

crlbmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

 crlbeq
 crlbmj
```

```{r sqrbl, fig.pos='H' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Blue", echo=FALSE, include=TRUE}

sqrbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbomj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbeq
sqrbomj
```

# ANNEX B  {-}

```{r heatmapdsquare,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Square orange, heterogenous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle \nß value 10" ,"Circle \nß value 5", "Circle \nß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_sq_or", "VALUE" = "val_sq_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_iv_square, y = beta_ie_square, fill = Segra)) + geom_tile() + facet_grid(rev_beta_iv_circle ~  beta_ie_circle + segregation, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs, beta_ie_circle = beta_ie_circle.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß value square") + ylab("ß ethnic square") + ggtitle("Square orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

```{r heatmapdcircle, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, heterogeneous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5","Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")
beta_iv_square.labs <- c("Square \nß value 0", "Square \nß value 5","Square \nß value 10")
names(beta_iv_square.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_cl_or", "VALUE" = "val_cl_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_ie_circle, y = beta_iv_circle, fill = Segra)) + geom_tile() + facet_grid(rev_beta_ie_square ~  beta_iv_square + segregation, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = beta_iv_square.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß ethnic circle") + ylab("ß value circle") + ggtitle("Circle orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

## Fig: \ref{fig:heatmapdsquare}
*  Fig: \ref{fig:heatmapdsquare} reports in heatmaps $\beta^{\square}_e$ on y-axis $\beta^{\square}_v$ on x-axis for square agents, as such on y-axis is the most important preference for the value group square, the triangle shape is due to the condition $\beta^{\square}_v \leq \beta^{\square}_e$.
* Faceted for the condition of the other group, focusin on 3 degree: total randomness $\beta^{\circ} = 0$, threshold between area of high and low randomness $\beta^{\circ} = 5$ and highest systematic component $\beta^{\circ} = 10$. For each condition ethnic and value segregation of squared population is calculated
* Lower area, where the other group shows total random behavior ($\beta^{\circ} = 0$) for both ethnic and value utility, value integration emerges with low $\beta^{\square}_v$ regardless the increase in $\beta^{\square}_e$. Higher value segregation occurs only for the highest levels of $\beta^{\square}_v$ and $\beta^{\square}_e$. Ethnic segregation increases along the diagonal where $\beta^{\square}_v = \beta^{\square}_e$. Lower ethnic segregation compared to the other conditions in the area of high ethnic utility $\beta^{\square}_e$
* Second row, first panel: for total random value behavior of square $\beta^{\square}_v = 0$, the ethnic segregation is linear to the increase of  $\beta^{\square}_e$, not showing the threshold difference as in Fig: \ref{fig:crlo}, this only with ethnic randomness of circle agents $\beta^{\circ}_e = 0$, but reach the critical point of $\beta^{\circ}_v = 5$ for value preferences. The sensitivity to change persists, but the sensitivity to chance decreases, because agents are split into two groups, for the same dimension that both share (ethnicity), one maximizes, the other not, and viceversa one doesn't care about value, while the other group maximizes value.
* My guess: circle agents would gather together in ethnically mixed neighborhoods (cfr Fig: \ref{fig:heatmapdcircle}, ethnic integration along $\beta^{\circ}_v$), which attract square ethnicity-oriented at increase of  $\beta^{\square}_e$, they not care about the value orientation of the neighborhood. Compared to the previous section, $\beta^{\square}_e$ applies to $50\%$ of the population, so that the reduced amount of square agents entering the neighborhood would be tolerated by circle, but not when the $\beta^{\square}_e$ increases and too square agents enter the neighborhood, which would explain why value segregation persists rather than value integration as would be expected.
* In the next panel where circle hold minimum threshold for both ethnic and value preferences $\beta^{\circ}_v = \beta^{\circ}_e = 5$, the demark between high randomness for $\beta^{\square}_e < 5$ and high utility $\beta^{\square}_e \geq 5$ returns and corresponds to value integration.
* value segregation increases homogeneously when circle population maximizes  $\beta^{\circ}_v = 10$

## Fig: \ref{fig:heatmapdsquare}
* Fig: \ref{fig:heatmapdsquare}, for cirlce agents with higher value preferences than ethnic preferences, ethnic integration emerges for area of ethnic randomness with $\beta^{\circ}_e < 5$
* Value segregation on the top right, where square agents show maximal ethnic and value utility  $\beta^{\square}_e = \beta^{\square}_v = 10$, full value segregation emerges for all conditions of $\beta$ levels, while in high ethnic utility area $\beta^{\square}_e \geq 5$ circle live in neighborhoods both value and ethnic homogeneous, for $\beta^{\square}_e <  5$ they live in ethnically mixed neighborhoods with circle of other ethnicity