### 2020-05-07 exp_eth_val.Rmd
presentation

---
output:
  bookdown::pdf_document2:
    toc: false
    keep_tex: true
    fig_caption: yes
title: "Observations"
# subtitle: "Target: BIGSSS Research Day"
author: Rocco Paolillo

# date: JU 
#language: en-GB
bibliography: "references.bib"
# always_allow_html: yes

# target: BIGSSS Research Day by September 15th, presented  on October 25th. 6000 words: around 12 pages.
# csl: apa.csl # by default: Chicago style
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
- \usepackage{amsmath}
- \usepackage{graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.lp="fig:"  ,warning = FALSE)

library(knitr)
library(ggplot2)
library(forcats)
library(reshape2)
library(tidyr)
library(dplyr)
library(reshape)
library(ggpubr)
library(magick)
library(gridExtra)
library(grid)

#theme_set(theme_pubr())

wd <- setwd("C:/Users/rocpa/OneDrive/Desktop/AF/")
dataone <- read.csv("experiments/80_50_onebeta.csv",sep = ";",skip = 1)
datatwo <- read.csv("experiments/two_beta.csv", sep = ";",skip = 1)

dataone$rev_circle_orange <- fct_rev(factor(dataone$circle_orange))
dataone$rev_circle_blue <- fct_rev(factor(dataone$circle_blue))

datatwo$rev_beta_iv_circle <- fct_rev(factor(datatwo$beta_iv_circle))
datatwo$rev_beta_ie_square <- fct_rev(factor(datatwo$beta_ie_square))

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")
 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")
 circle_orange.labs <-  c("20% Circle \nOrange", "50% Circle \nOrange","80% Circle \nOrange")
names(circle_orange.labs) <- c("20", "50","80")

```
\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}

# Recap

- Research question: How the interplay of ethnic and value homophily preferences can explain difference in the degree of integration for the members of the same ethnic group (hybrid segregation) for both ethnic and value segregation.
- Focus: different weight on ethnic homogeneity ($\beta_e$) and value homogeneity ($\beta_v$), heterogeneity among agents, relative group size for both dimensions \par

Utility of agents in Schelling relocation

$$
U= \beta_eX_v + \beta_vX_v + \epsilon
$$

```{r thheory, fig.pos="h", out.width="50%",out.width='\\textwidth',fig.align='left',fig.show="hold", fig.cap="Agents distinction (subgroups) and utility function", echo=FALSE, include=TRUE }

knitr::include_graphics("recapp.png")
```

# First Section (Tab: \ref{tab:First})

* Aim:
    + Effect of different homophily preferences: direct effect of $\beta_e$ on ethnic segregation, and direct effect of beta value $\beta_v$ on value segregation; how $\beta_v$ influences effect of  $\beta_e$ on ethnic segregation and how $\beta_e$ influences effect of $\beta_v$ on value segregation
* Focus: relative group size for ethnicity and value distribution within each ethnic group, considering how they directly influence probability to sort with similars; interaction with levels of $\beta_e$,$\beta_v$
* Assumption: 
    + not heterogeneity: all agents hold both ethnic preference (same color) and value preference (same shape) depending on level of global $\beta_e$ and $\beta_v$
    + equal ethnic size (blue/orange: $0.50$) vs majority/minority (blue/orange: $0.80$)
    + value distribution within each ethnic group: proportion circle/square, moving from minority circle to majority cirlce:  $0.20$,$0.50$,$0.80$. It can be seen as moving from more intolerant to  more intolerant society when value-dependent heterogeneous preferences are included. In this section, there is not heterogeneous preferences
* Figures:
    + Fig: \ref{fig:baselineone} represents how results are presented through the space of the model. I identified $\beta =5$ for both ethnic and value preferences as threshold between area of high randomness and low randomness. Results can be explained as within the space resulting from area of high randomness and low randomness (i.e. high systematic component) for ethnic and value utility.
    + Relative group sizes: the bold frame in Fig: \ref{fig:global} would represent the baseline, where each subgroup in Fig: \ref{fig:thheory} represents the $25 \%$ of the population and has probability $0.25$ to meet an agent of same ethnicity and same value (baseline: blue/orange = $0.5$ and within each ethnic group cirlce/square = $0.5$)
    + Relative group sizes (value orientation): in each figure from Fig: \ref{fig:global} to Fig: \ref{fig:crlblmajority} along the diagonal, moves from minority circle (low corner) in the population to majority circle (upper corner)
    + x-axis = ethnic weight $\beta_e$, y-axis = value weight $\beta_v$

\begin{table}[H]
\begin{tabular}{llll}
\midrule
 Static agent variable                      & Range         &      &                \\ \midrule
 Ethnicity  (\textit{color})         & Blue, Orange        \\
 Value orientation (\textit{shape}) & Square, Circle &  &  \\
 \midrule

 Parameters kept constant    & Slider & Range   & Experiments  \\ 
 \midrule
 Population density    & density                 & $[0,0.99]$   & $0.7$          \\
 Peak ethnic utility & similar\_wanted\_ethnic & $[0,1]$ & $1$ (linear function) \\
 Peak value utility & similar\_wanted\_value & $[0,1]$ & $1$ (linear function) \\
 \midrule
 \textbf{Parameters modified in experiment} & \textbf{Slider} & \textbf{Range}   & \textbf{Experiments}  \\ 
 \midrule
 Ethnic group ratio (local\/minority)  & fraction\_blue &  $[0,1]$      & $0.5,0.8$  \\
 value ratio  blue population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 value ratio  orange population & circle\_blue & $[0,1]$      & $0.20,0.5,0.8$  \\
 Determinism ethnic utility ($\beta_e$) & beta-ie for all agents & $[0,10]$ & $[0,10]$  \\
 Determinism value utility ($\beta_v$) & beta-iv for all agents & $[0,10]$ & $[0,10]$ \\
 \midrule
 Global output measures    & Computation &   Range & \\ 
 \midrule
 Ethnic segregation  & same color neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
 Value segregation  & same shape neighborhood $\%$  & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
   \midrule
Number of Runs   & 1 & Numer of steps & 1000 \\ 
 \midrule
\end{tabular}
 \caption{First Section} 
 \label{tab:First}
\end{table}

```{r baselineone,fig.pos='H',out.width="96%",fig.show = "hold",fig.align='center', fig.cap="Space of model", echo=FALSE, include=TRUE  }

knitr::include_graphics("baseline_1.png")

```

## Global value and  ethnic segregation (Fig: \ref{fig:global}, Fig: \ref{fig:global-maj})

* Fig: \ref{fig:global}: equal ethnic size, Fig: \ref{fig:global-maj}: majority/minority condition
* Ethnic segregation increases along the axis of $\beta_e$, while value segregation symmetrically along the axis of $\beta_v$. In both cases $\beta = 5$ seems a critical limit between high randomness and low randomness (high systematic utility). Consequently, area in the space $\beta < 5$ is closer  to ethnic and value integration, due to random relocation, area $\beta >= 5$ shows complete segregation because of the systematic component
* While ethnic and value segregation linearly increase for the high randomness area below  $\beta = 5$, segregation for both dimensions for $\beta > 5$ seems homogeneous at high level
* The limit of $\beta = 5$ to me can relate to "sensitivity to change" in @van2009neighborhood: since the shape of utility is linearly increasing, each difference in neighborhood composition corresponds to a change in utility. The product $\beta * \Delta Utility$ is affected by randomness only for very low levels of $\beta$, with $\beta =5$ it becomes too strong: agents are too sensitive to the systematic utility, moving to the neighborhood with the higher utility, affecting its composition, which will increase the utility difference and attractiveness of the neighborhood to agents in next step, falling shortly into complete segregation. Difference for higher levels of $\beta = 5$ is negligible.
* At $\beta = 0$ for each dimnension, increase in the other $\beta$ ($\beta_v$ on $\beta_e$ and $\beta_e$ on $\beta_v$) does not produce effects, apart in the area of high randomness when one value group represents the minority in one ethnicity and the majority on the other (upper left and lower right panel). Affected  by relocation of majority value.
* Condition of ethnic majority/minority shows basically the same patterns, slightly higher ethnic segregation for the area of high randomness, which is direct effect of unequal group sizes.

```{r global, fig.height=8.8, fig.pos="H", include=TRUE, fig.cap="Global Segregation with ethnic equal group size, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

 circle_local.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_local.labs) <- c("20", "50","80")
 circle_minority.labs <-  c("20% Circle \nOrange","50% Circle \nOrange","80% Circle \nOrange")
names(circle_minority.labs) <- c("20", "50","80")

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")
 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

eth_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(rev_circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  rev_circle_orange = rev_circle_orange.labs ), switch  = "y")  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10), position = "right")   + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
   geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

val_eq <- ggplot(data=subset(dataone,fraction_blue == 50 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(rev_circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  rev_circle_orange = rev_circle_orange.labs ), switch  = "y")  + 
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10), position = "right")  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 50 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

glb <- annotate_figure(ggarrange(eth_eq , ggarrange(val_eq), nrow = 2), top = text_grob("Global Segregation, Blue:Orange = 0.5:0.5",face = "bold", size = 12))

glb

```

```{r global-maj, fig.height=8.8, fig.pos="H",  include=TRUE, fig.cap="Global Segregation with majority/minority ethnicity, baseline in boldline", echo=FALSE}

# gen_eq <- filter(dataone, fraction_blue == 50 & step == 1000)
# gen_mj <- filter(dataone, fraction_blue == 80 & step == 1000)

 circle_local.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_local.labs) <- c("20", "50","80")
 circle_minority.labs <-  c("20% Circle \nOrange","50% Circle \nOrange","80% Circle \nOrange")
names(circle_minority.labs) <- c("20", "50","80")

# eth_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= eth_seg))
# val_mj <- ggplot(gen_mj, aes( beta_iv, beta_ie,fill= val_seg))
#

maj_eth <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= eth_seg)) + geom_tile() +   facet_grid(rev_circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  rev_circle_orange = rev_circle_orange.labs ), switch  = "y")  +
  scale_fill_distiller(name = "ETHNIC",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
  scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10), position = "right") + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_val <- ggplot(data=subset(dataone,fraction_blue == 80 & step == 1000), aes( beta_ie, beta_iv,fill= val_seg)) + geom_tile() +   facet_grid(rev_circle_orange  ~  circle_blue,  space = "free_x", labeller = labeller(circle_blue = circle_local.labs,  rev_circle_orange = rev_circle_orange.labs ), switch  = "y")  +
  scale_fill_distiller(name = "VALUE",palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) + 
 scale_x_discrete(name = "ß ethnic", limits=seq(0,10 ,by = 1)) + 
  scale_y_discrete(name = "ß value", limits=c(0,3,5,7,10), position = "right")  + 
  theme_bw() + theme(strip.text.x = element_text(), legend.position = "right", panel.grid.minor.x = element_blank(), plot.title=element_blank(),  legend.margin=margin(0,0,0,0)) +
  geom_rect(data=subset(dataone,fraction_blue == 80 & step == 1000 & circle_orange == 50 & circle_blue == 50),aes(xmin = -1,xmax = 12, ymin = -1, ymax = 11),colour="black",size =2,alpha = 0)

maj_glb <- annotate_figure(ggarrange(maj_eth , ggarrange(maj_val), nrow = 2), top = text_grob("Global Segregation, Blue:Orange = 0.8:0.2",face = "bold", size = 12))
maj_glb

```

## Subgroup behavior (Fig: \ref{fig:crlo} to Fig: \ref{fig:crlblmajority})

* Circle orange is used as model for equal ethnic size and ethnic group in the minority condition. Circle  blue as model for ethnic group in majority condition. The observed behavior can be applied symmetrically to the other subgroups imposing same conditions, I do this in ANNEX A rotating the axes.
* 3 outcomes are evident, rather than 2 as in traditional assessment of Schelling: integration (filler white  0.5), segregation (filler red above 0.5), and assimilation (filler blue close to 0, only members outgroup). This can relate as measure of *hybrid segregation*

### Equal ethnic group size (Fig: \ref{fig:crlo})

* Baseline reproduces the same as global observations. Ethnic segregation doesn't change along the diagonal; value segregation does, as effect of different value distribution (minority circle in the low left; majority circle in the upper right)
* Top left corner, circle is majority value in the own ethnic group (orange), minority in the other group (blue): integration occurs in the high randomnessfor both preferences (low left corner); ethnic segregation in high random ethnic area increases with increase of $\beta_v$; increase also of value segregation for increase in $\beta_e$: Both results  are due to the majority value circle increasing both value and  ethnic preferences, in condition of equal ethnic size.
* Top row: proceding on x-axis with increase in percentage similar value in the other ethnic group: for area of ethnic randomness, ethnic integration more evident over ethnic segregation at increase of $\beta_v$, because more equal chance to satisfy value preferences with agents of both ethnic groups
* Bottom left corner, circle value is minority value in both groups: value assimilation to square happens for all levels of $\beta_e$, as direct consequence of minority position. Ethnic integration emerges for high ethnic randomness ($\beta_e=5$)  also for increase in $\beta_v$, because agents need to cluster with value similars of the other ethnic group, which still are a minority. Value segregation results constantly high for all area  $\beta_v \geq 5$. Still, looking at overlap of ethnic and value space, moving from high ethnic randomness to low ethnic randomness area, ethnic integration shifts to ethnic segregation. This means that agents still manage to live in neighborhoods value homogeneous, but only with members of the own ethnic group. Maybe good to include density observation here, and compare with ACS.
* Bottom row: proceding on x-axis with increase of circle in the other ethnic group: value assimilation for $\beta_v < 5$ narrows down to high $\beta_e$, since for random ethnic behavior there is more chance to find value similar in the other group being equally represented. When value preference increases, ethnic assimilation becomes more evident since circle have more chance to find similar value, though of the other ethnic group, which they don't care about.
* Middle row (circle are 50% value in ethnic group): transition in between observations above, not  different patters

```{r crlo,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, equal ethnic size", echo=FALSE, include=TRUE  }

dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

```

### Ethnic minority (Fig: \ref{fig:crlominority})

* Trends are generally similar to Fig: \ref{fig:crlo}
* In the baseline, ethnic assimilation emerges instead of ethnic integration, becuase of the ethnic minority of the group.
* Up left corner, where circle represent the majority value of ethnic  minority and the minority value of ethnic majority, compared to Fig: \ref{fig:crlo} ethnic assimilation rather than integration emerges for the area of high randomness for both dimensions, and ethnic integration rather than segregation for the increase of $\beta_v$ and low values of $\beta_e$. While for areas of high randomness agents fall into assimilation because of numeric inferiority, when they increase value utility $\beta_v$ they want to relocate close to other circle agents, regardless of their ethnicity. Since in this condition circle represent the minority value of the majority ethnic population (the opposite of circle orange in ethnic minority), the combination allows for the increase in value preference to make up for the ethnic numeric inferiority, so that both circle minority and circle majority end up in neighborhoods ethnically integrated and value circle homogeneous (cfr Fig: \ref{fig:crlbl}, right down corner).
* Low row: the effect of increase in $\beta_e$ on value assimilation is similar as Fig: \ref{fig:crlo}, since circle here represents the minority in the own group. Increase of $\beta_v$ shows instead higher assimilation, because circle minority has more chance to relocate close to similar circle value of the other ethnic group, proportionally to value circle being the majority value in the majority ethnic group. Interestingly, given the area of high ethnic utility for $\beta_e \geq 5$, at the increase of $\beta_v$ less homogenous ethnic segregation emerges compared to the equal ethnic size.
* The last two  points are noteworthy to me: the assumption that similar values can foster ethnic integration over ethnic segregation can work provided that an adjustement between distribution of values in both ethnic group and preferences for ethnic or value utility are  taken into account.

```{r crlominority,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, ethnic minority", echo=FALSE, include=TRUE  }

dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_or","Ethnic" = "eth_cl_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(rev_circle_orange ~ circle_blue + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

```

### Ethnic majority (Fig: \ref{fig:crlblmajority})
* Low row where circle are the minority in the ethnic majority blue, show the same pattern for the circle in minority and equal ethnic size: increasing in ethnic utility implies value assimilation to square  members of the same ethnic group, because of value numeric minority
* Value integration occurs in the upper row ($50 \% $ circle in blue majority ethnic) for random area of low $\beta_v$, because eqaul chance to relocate close to either circle or square, but in both cases of same ethnic group
* Compared to the baseline in Fig: \ref{fig:crlo}, ethnic segregation emerges also in area of high randomness ($\beta_e < 5$), because of numeric superiority, also for the upper side of the diagonal
* Low right corner, where circle are the minority in the blue ethnicity, but majority in the ethnic minority, ethnic integration emerges for high values of $\beta_v$, despite ethnic minority of orange. Same as in Fig: \ref{fig:crlominority}

```{r crlblmajority, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="Circle blue, ethnic majority", echo=FALSE, include=TRUE}

  rev_circle_blue.labs <- c("80% Circle \nBlue" ,"50% Circle \nBlue","20% Circle \nBlue")
names(rev_circle_blue.labs) <- c("80", "50","20")

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

```

# Second Section (heterogeneous preferences for tolerant and intolerants)

* Aim:
    + Test the effect of heterogeneous preferences for two value groups, i.e. circle value-oriented care more about value than ethnic neighborhood composition, square ethnicity-oriented care more about ethnic than value composition
    + This would relate to the difference between tolerant (circle) and intolerant (square) as in the ACS paper and that we discussed
    + Observe how heterogeneity in agents' preferences along two value oriented groups will affect results in the previous section, through their mutual adaptation
    + Interaction between heterogeneous preferences and relative group sizes for both dimensions (here only equal ethnic and value group)
* Assumption: 
    + Agents hold both ethnic and value preferences, but $\beta$ for each dimension depends on their value orientation, either circle or square. Not difference due to ethnic membership are taken into consideration
    + Given the linear function, utility increases with value or ethnic homogeneity of neighborhood, but determinism component $\beta_e$ and $\beta_v$ are differently distributed to value groups: ethnic preferences of circle cannot be higher than value preferences ($\beta^{\circ}_e \leq \beta^{\circ}_v$) and value preferences of square cannot be higher than thier value preferences ($\beta^{\square}_v \leq \beta^{\square}_e$)
    + Ideally also condition of ethnic majority/minority and circle majority/minority should be included. Here for matter of time it is reduced to equal ethnic group size (blue/orange $0.50$) and for each ethnic group equal value group (circle/square $0.50$). So, each of the four subgroup in Fig: \ref{fig:thheory} represents the $25 \%$ of the global agents' population
* Figures:
    + I compare the ethnic and value segregation of both circle and square agents
    + Focus on the difference between ethnic and value segregation (being close to agents sharing both ethnicity and value)
    + To reduce observations given the numbers of parameters, I jump on 3 critical points individualized in the previous section: $\beta = 0$: total randomness, $\beta = 5$: min. determinism, $\beta = 10$: max determinism
    + For each value group, in the baseline I put on x-axis the difference between the highest reference $\beta$ of that group and the other one, i.e. $\beta_e - \beta_v$ for square, $\beta_v - \beta_e$ for circle. Considering the critical point, I repeat observation for reference $\beta = 5$: min. determinism, and $\beta = 10$: max determinism, $\beta = 0$ is included in bot. Thus, along the x-axis the group becomes more sensitive to the reference dimension, ethnicity for square group, value for circle group. I compare how the baseline condition changes for interaction of value and ethnic preferences of the other group faceting the 3 critical points $\beta = 0; \beta = 5;\beta = 10$

\begin{table}[H]
\begin{tabular}{llll}
\midrule
 Static agent variable                      & Range         &      &                \\ \midrule
 Ethnicity  (\textit{color})         & Blue, Orange        \\
 Value orientation (\textit{shape}) & Square, Circle &  &  \\
 \midrule

 Parameters kept constant    & Slider & Range   & Experiments  \\ 
 \midrule
 Population density    & density                 & $[0,0.99]$   & $0.7$          \\
 Peak ethnic utility & similar\_wanted\_ethnic & $[0,1]$ & $1$ (linear function) \\
 Peak value utility & similar\_wanted\_value & $[0,1]$ & $1$ (linear function) \\
 \midrule
 \textbf{Parameters modified in experiment} & \textbf{Slider} & \textbf{Range}   & \textbf{Experiments}  \\ 
 \midrule
 Ethnic group ratio (local\/minority)  & fraction\_blue &  $[0,1]$      & $0.5$  \\
 value ratio  blue population & circle\_blue & $[0,1]$      & $0.5$  \\
 value ratio  orange population & circle\_blue & $[0,1]$      & $0.5$  \\
 Determinism ethnic utility square ($\beta^{\square}_e$) & beta-ie-sqr & $[0,10]$ & $[0,10]$  \\
 Determinism value utility square ($\beta^{\square}_v$) & beta-iv-sqr  & $[0,10]$ & $[0,10], if \beta^{\square}_v \leq \beta^{\square}_e$ \\
 Determinism value utility circle ($\beta^{\circ}_v$) & beta-iv-crl  & $[0,10]$ & $[0,10]$  \\
 Determinism ethnic utility circle ($\beta^{\circ}_e$) & beta-ie-crl & $[0,10]$ & $[0,10], if \beta^{\circ}_e \leq \beta^{\circ}_v$  \\
 \midrule
 Global output measures    & Computation &   Range & \\ 
 \midrule
 Ethnic segregation  & same color neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
 Value segregation  & same shape neighborhood $\%$ & $0,\stackrel{+0.1}{\dots}, 1$ & global and subgroups\\
   \midrule
Number of Runs   & 1 & Number of steps & 1000 \\ 
 \midrule
\end{tabular}
 \caption{Second section: heterogeneous preferences} 
 \label{tab:Second}
\end{table}

## Comparison different value orientation
* Fig: \ref{fig:betatwoethnic}: x-axis $\beta^{\square}_e - \beta^{\square}_v$, on the top panel with $\beta^{\square}_e = 5$ (area of high randomness), on bottom panel  $\beta^{\square}_e = 10$ (area of low randomness). X-axis = 0,  ethnic and value utility have equal weight, along the x-axis ethnic weight prevails over value weight. the lower the distance between ethnic and value segregation, the more agents live close to others similar for both ethnicity and value.
* Viewing at lower left panel, where behavior of circle is totally random, not only the distance between ethnic and value segregation of square is lower when $\beta^{\square}_e = \beta^{\square}_v$ (x-axis 0), but also higher for both dimensions, progressing towards x-axis (ethnic becomes higher than 10), value segregation decreases, with more linear slope for high determinism area (bottom panel).
* It persists that each $\beta$ rules the reference dimension, but the random behavior of the other group decreases the segregation expected for that level of $\beta$, higher determinism of the other group increases  the level of segregation for that dimension. Demonstration, moving from $\beta^{\circ}_v = 5$ to  $\beta^{\circ}_v = 10$ 
* As $\beta^{\square}_e$ becomes higher than  $\beta^{\square}_v$, moving on the x-axis, ethnic segregation persists constantly, while value segregation diminishes.
* The threshold of $\beta = 5$ seen in the previous section is somehow replicated: in the top panel, max value and ethnic segregation are equal 1 when both  $\beta^{\square}_e = \beta^{\square}_v = 0$, on the x-axis 0, and value segregation declines once $\beta^{\square}_v$ comes close to 0. Looking at panel below (high determinism), the decline of value segregation occurs equally when $\beta^{\square}_v$ declines from 5 to 0 (on the right of red line 10-5:5 and so on)
* Even though the direct effect of $\beta_v$ is assumed and same patterns are observed from low determinism (top panel) and high determinism (bottom panel), the high determinism of  $\beta^{\square}_e$ is associated with very slight increase in the value segregation of square, also for same level of $\beta^{\square}_v = 0$, though negligeable
* The distance between ethnic and value segregation is higher in the middle panel, where circle hold same min utility preferences $\beta^{\circ}_e = \beta^{\circ}_v = 5$, and also square increase in ethnic over value preference (x-axis = 5). Basically the only random component is the value orientation of square agents, which has the same effect on both circle and square population patterns
* Moving on the diagonal, where $\beta^{\circ}_e = \beta^{\circ}_v = 10$, the distance between ethnic and value decreases, meaning agents are more likely to live close to agents of both same ethnicity and value: the increase in utility at max level of circle ($\beta^{\circ}_e = \beta^{\circ}_v = 10$) compensates for random value behavior of square, since circle will avoid neighborhoods with square 
* Fig: \ref{fig:betatwovalue} shows the similar trends, focusing on value segregation over ethnic segregation for the circle population. Nevertheless, in the middle panel ($\beta^{\square}_e = \beta^{\square}_v = 5$), I would expect symmetric scenario than in Fig: \ref{fig:betatwoethnic}, with ethnic segregation of circle being same as value segregatiion than square. Why here circle are completed ethnically integrared? There is a difference between the two dimensions? Think about.

```{r betatwoethnic,fig.pos='H',fig.show = "hold",fig.align='center',fig.cap="Difference segregation, x-axis: difference ethnic and value preference square, faceting conditions: preferences of circle", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle ß value 10" ,"Circle ß value 5", "Circle ß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

#  v_circle.labs <- c("ß value circle 0", "ß value circle 5", "ß value circle 10")
# names(v_circle.labs) <- c("0", "5","10")
e_circle.labs <- c("Circle ß ethnic 0", "Circle ß ethnic 5", "Circle ß ethnic 10")
names(e_circle.labs) <- c("0", "5","10")

ie_10 <- circle_frame %>%  filter(beta_ie_square == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß ethnic square 10 - ß value square", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5 <- circle_frame %>%  filter(beta_ie_square == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority"))) %>%
  ggplot(aes(x = (beta_ie_square - beta_iv_square), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_iv_circle ~  beta_ie_circle, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs,  beta_ie_circle = e_circle.labs )) +
  scale_x_discrete(name = "ß ethnic square 5 - ß value square", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

ie_5
ie_10
```

```{r betatwovalue,  fig.pos="H",fig.show = "hold", fig.align='center',fig.show="hold", fig.cap="Difference segregation between", echo=FALSE, include=TRUE  }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5", "Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")

e_square.labs <- c("ß ethnic \nsquare 0", "ß ethnic \nsquare 5", "ß ethnic \nsquare 10")
names(e_square.labs) <- c("0", "5","10")
v_square.labs <- c("ß value square 0", "ß value square 5", "ß value square 10")
names(v_square.labs) <- c("0", "5","10")

iv_10 <- circle_frame %>%  filter(beta_iv_circle == 10) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  geom_vline(xintercept = 5, color = "red") +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,10, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5 <- circle_frame %>%  filter(beta_iv_circle == 5) %>% gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or,eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ethnic segregation" = c("eth_sq_or","eth_cl_or"), "value segregation" = c("val_sq_or","val_cl_or")), "ethnic segregation"), `group` = fct_relevel(fct_collapse(subgroup, "square minority" = c("eth_sq_or","val_sq_or"), "circle minority" = c("eth_cl_or","val_cl_or"), "square minority")))  %>%
  ggplot(aes(x = (beta_iv_circle - beta_ie_circle), y = Segra, color = `segregation`, shape = `group`)) + geom_line() + geom_point( size = 2) + facet_grid(rev_beta_ie_square ~  beta_iv_square, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = v_square.labs))  +
  scale_x_discrete(name = "ß value circle 10 - ß ethnic circle", limits = seq(0,5, by = 1)) +  ylab("Segregation") +
  scale_color_manual(values = c("dark violet","dark green"), guide = guide_legend(title="")) +
  scale_shape_manual(values = c(16,15), guide = guide_legend(title="")) +
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

iv_5
iv_10

```

# ANNEX A {-}

```{r sqro, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Orange", echo=FALSE, include=TRUE  }

sqroeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroqmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_or","Ethnic" = "eth_sq_or"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_orange ~ rev_circle_blue + segregation, labeller = labeller(circle_orange = circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square orange, ethnic minority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqroeq
sqroqmj
```

```{r crlbl,fig.pos='H' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Circle Blue", echo=FALSE, include=TRUE}

 rev_circle_orange.labs <-  c("80% Circle \nOrange", "50% Circle \nOrange","20% Circle \nOrange")
names(rev_circle_orange.labs) <- c("80", "50","20")

 circle_blue.labs <- c("20% Circle \nBlue", "50% Circle \nBlue","80% Circle \nBlue")
names(circle_blue.labs) <- c("20", "50","80")

crlbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

crlbmj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_cl_bl,val_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_cl_bl","Ethnic" = "eth_cl_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid( rev_circle_blue  ~ circle_orange  + segregation, labeller = labeller(circle_orange =circle_orange.labs, rev_circle_blue = rev_circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Circle blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

#grid.arrange(grob(sqroeq),grob(sqromj), ncol =  1, common.legend = TRUE,legend = "bottom")

 crlbeq
 crlbmj
```

```{r sqrbl, fig.pos='H' ,out.width='\\textwidth',fig.show = "hold",fig.align='center',fig.show="hold", fig.cap="First Section, Square Blue", echo=FALSE, include=TRUE}

sqrbeq <- dataone %>%  filter(fraction_blue == 50) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, equal ethnic size") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbomj <- dataone %>%  filter(fraction_blue == 80) %>%  gather(key = subgroup, value = Segra, eth_sq_bl,val_sq_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Value" = "val_sq_bl","Ethnic" = "eth_sq_bl"), "Value"))  %>% 
  ggplot(aes(x = beta_ie,y = beta_iv, fill = Segra)) + geom_tile() + 
  facet_grid(circle_blue ~  rev_circle_orange  + segregation, labeller = labeller(rev_circle_orange = rev_circle_orange.labs, circle_blue = circle_blue.labs), switch  = "y") +
  scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5), position = "right") +
  xlab("ß ethnic") + ylab("ß value") + ggtitle("Square blue, ethnic majority") + 
  theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0), 
                      panel.spacing.x  = unit(0.2, "lines"),panel.spacing.y  = unit(0.1, "lines") )

sqrbeq
sqrbomj
```

# ANNEX B  {-}

```{r heatmapdsquare,fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Square orange, heterogenous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_iv_circle == 0 |beta_iv_circle == 5 |beta_iv_circle == 10 ) %>% filter(beta_ie_circle == 0 | beta_ie_circle == 5 | beta_ie_circle == 10)

rev_beta_iv_circle.labs <- c("Circle \nß value 10" ,"Circle \nß value 5", "Circle \nß value 0")
names(rev_beta_iv_circle.labs) <- c("10", "5","0")
beta_ie_circle.labs <- c("Circle \nß ethnic 0", "Circle \nß ethnic 5","Circle \nß ethnic 10")
names(beta_ie_circle.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_sq_or,val_sq_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_sq_or", "VALUE" = "val_sq_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_iv_square, y = beta_ie_square, fill = Segra)) + geom_tile() + facet_grid(rev_beta_iv_circle ~  beta_ie_circle + segregation, labeller = labeller(rev_beta_iv_circle = rev_beta_iv_circle.labs, beta_ie_circle = beta_ie_circle.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß value square") + ylab("ß ethnic square") + ggtitle("Square orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

```{r heatmapdcircle, fig.pos='H',out.width='\\textwidth',fig.show = "hold",fig.align='center', fig.cap="Circle orange, heterogeneous preferences", echo=FALSE, include=TRUE }

circle_frame <- datatwo %>% filter(step == 1000) %>% filter(beta_ie_square == 0 |beta_ie_square == 5 |beta_ie_square == 10 ) %>% filter(beta_iv_square == 0 | beta_iv_square == 5 | beta_iv_square == 10)

rev_beta_ie_square.labs <- c("Square \nß ethnic 10" ,"Square \nß ethnic 5","Square \nß ethnic 0")
names(rev_beta_ie_square.labs) <- c("10", "5","0")
beta_iv_square.labs <- c("Square \nß value 0", "Square \nß value 5","Square \nß value 10")
names(beta_iv_square.labs) <- c("0", "5","10")

circle_frame %>%   gather(key = subgroup, value = Segra, eth_cl_or,val_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"ETHNIC" = "eth_cl_or", "VALUE" = "val_cl_or"), "ethnic segregation")) %>%
  ggplot(aes(x = beta_ie_circle, y = beta_iv_circle, fill = Segra)) + geom_tile() + facet_grid(rev_beta_ie_square ~  beta_iv_square + segregation, labeller = labeller(rev_beta_ie_square = rev_beta_ie_square.labs, beta_iv_square = beta_iv_square.labs)) +
   scale_fill_distiller(palette = "RdBu", limits=c(0,1), breaks=c(0,0.2,0.5,0.8,1)) +
  scale_x_discrete( limits=c(0,10, by = 5)) +
  scale_y_discrete( limits=c(0,10, by = 5)) +
 xlab("ß ethnic circle") + ylab("ß value circle") + ggtitle("Circle orange, baseline") + 
    theme_bw() +  theme(legend.position = "bottom", legend.title = element_blank(), panel.grid.minor.x = element_blank(), plot.title=element_text(hjust = 0.5),  legend.margin=margin(0,0,0,0))

```

## Fig: \ref{fig:heatmapdsquare}
*  Fig: \ref{fig:heatmapdsquare} reports in heatmaps $\beta^{\square}_e$ on y-axis $\beta^{\square}_v$ on x-axis for square agents, as such on y-axis is the most important preference for the value group square, the triangle shape is due to the condition $\beta^{\square}_v \leq \beta^{\square}_e$.
* Faceted for the condition of the other group, focusin on 3 degree: total randomness $\beta^{\circ} = 0$, threshold between area of high and low randomness $\beta^{\circ} = 5$ and highest systematic component $\beta^{\circ} = 10$. For each condition ethnic and value segregation of squared population is calculated
* Lower area, where the other group shows total random behavior ($\beta^{\circ} = 0$) for both ethnic and value utility, value integration emerges with low $\beta^{\square}_v$ regardless the increase in $\beta^{\square}_e$. Higher value segregation occurs only for the highest levels of $\beta^{\square}_v$ and $\beta^{\square}_e$. Ethnic segregation increases along the diagonal where $\beta^{\square}_v = \beta^{\square}_e$. Lower ethnic segregation compared to the other conditions in the area of high ethnic utility $\beta^{\square}_e$
* Second row, first panel: for total random value behavior of square $\beta^{\square}_v = 0$, the ethnic segregation is linear to the increase of  $\beta^{\square}_e$, not showing the threshold difference as in Fig: \ref{fig:crlo}, this only with ethnic randomness of circle agents $\beta^{\circ}_e = 0$, but reach the critical point of $\beta^{\circ}_v = 5$ for value preferences. The sensitivity to change persists, but the sensitivity to chance decreases, because agents are split into two groups, for the same dimension that both share (ethnicity), one maximizes, the other not, and viceversa one doesn't care about value, while the other group maximizes value.
* My guess: circle agents would gather together in ethnically mixed neighborhoods (cfr Fig: \ref{fig:heatmapdcircle}, ethnic integration along $\beta^{\circ}_v$), which attract square ethnicity-oriented at increase of  $\beta^{\square}_e$, they not care about the value orientation of the neighborhood. Compared to the previous section, $\beta^{\square}_e$ applies to $50\%$ of the population, so that the reduced amount of square agents entering the neighborhood would be tolerated by circle, but not when the $\beta^{\square}_e$ increases and too square agents enter the neighborhood, which would explain why value segregation persists rather than value integration as would be expected.
* In the next panel where circle hold minimum threshold for both ethnic and value preferences $\beta^{\circ}_v = \beta^{\circ}_e = 5$, the demark between high randomness for $\beta^{\square}_e < 5$ and high utility $\beta^{\square}_e \geq 5$ returns and corresponds to value integration.
* value segregation increases homogeneously when circle population maximizes  $\beta^{\circ}_v = 10$

## Fig: \ref{fig:heatmapdsquare}
* Fig: \ref{fig:heatmapdsquare}, for cirlce agents with higher value preferences than ethnic preferences, ethnic integration emerges for area of ethnic randomness with $\beta^{\circ}_e < 5$
* Value segregation on the top right, where square agents show maximal ethnic and value utility  $\beta^{\square}_e = \beta^{\square}_v = 10$, full value segregation emerges for all conditions of $\beta$ levels, while in high ethnic utility area $\beta^{\square}_e \geq 5$ circle live in neighborhoods both value and ethnic homogeneous, for $\beta^{\square}_e <  5$ they live in ethnically mixed neighborhoods with circle of other ethnicity

# References {-}