### 2021-05-25 additional.Rmd
test 3D plot

---
title: "additional3D"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/rocpa/OneDrive/Desktop/AF/ethnic-value_multinomial-master/additional/")

library(dplyr)
library(ggplot2)
library(plotly)
library(scatterplot3d)
library(reshape)
library(reshape2)
library(tidyr)
library(forcats)
library(plot3D)

valueminority.labs <- c("ß value liberals minority: 0", "ß value liberals minority: 1", "ß value liberals minority: 7")
names(valueminority.labs) <- c("0","1","7")
valuemajority.labs <- c("ß value liberals majority: 0", "ß value liberals majority: 1", "ß value liberals majority: 7")
names(valuemajority.labs) <- c("0","1","7")
ethnicminority.labs <- c("ß ethnic liberals minority: 0", "ß ethnic liberals minority: 1", "ß ethnic liberals minority: 7")
names(ethnicminority.labs) <- c("0","1","7")
ethnicmajority.labs <- c("ß ethnic liberals majority: 0", "ß ethnic liberals majority: 1", "ß ethnic liberals majority: 7")
names(ethnicmajority.labs) <- c("0","1","7")

mech <- read.csv("mech_lib.csv",sep=",")
excl <- names(mech) %in% c("run", "density","step")
mech <- mech[!excl] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

df_sec <- read.csv("df_secondary.csv",sep=",")
exclet <- names(df_sec) %in% c("run", "density","step")

df_sec <- df_sec[!exclet] %>% group_by(ID,RND,eth_con_loc,eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

 df <- read.csv("3Dlowdom.csv",sep = ",",skip = 6)
df2 <- read.csv("3Dhighdom.csv",sep = ",",skip = 6)
leg <- read.csv("legend_sa_asym.csv",sep=";")[,1]
df_fin <- rbind(df,df2)
 names(df_fin)<- leg

# 
# df_t <- rbind(df_3D,df_fin)
# 
# 
# write.csv(df_t,"3D.csv",row.names = F)

df_3D <- read.csv("3D.csv",sep=",")
excl3D <- names(df_3D) %in% c("run", "density","step")
df_3D <- df_3D[!excl3D] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

```

```{r asym_baseline, fig.cap="Fig. 8 orig", echo = FALSE}

asym_dom <- read.csv("df_asym_dom.csv",sep = ",")

excl <- names(asym_dom) %in% c("run", "density","step")
df <- asym_dom[!excl] %>% group_by(dominant,circle_orange,circle_blue,fraction_blue,secondary) %>% summarise_all(mean)

df <- df %>% filter(df$dominant <= 15) 
df <- df %>% filter(df$fraction_blue==50|df$fraction_blue==60|df$fraction_blue==80)

majcons <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_sq_bl, colour = vl_sq_bl),shape = "circle")  + geom_line(aes(x=dominant,y=dv_sq_bl, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_sq_bl, colour = et_sq_bl),shape="square") + geom_line(aes(x=dominant,y=de_sq_bl, linetype = as.factor(fraction_blue)))+ ggtitle("majcons"))

majlib <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_cl_bl, colour = vl_cl_bl),shape = "circle")  + geom_line(aes(x=dominant,y=dv_cl_bl, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_cl_bl, colour = et_cl_bl),shape="square") + geom_line(aes(x=dominant,y=de_cl_bl, linetype = as.factor(fraction_blue))) + ggtitle("majlib"))

mincons <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_sq_or, colour = vl_sq_or),shape = "circle")  + geom_line(aes(x=dominant,y=dv_sq_or, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_sq_or, colour = et_sq_or),shape="square") + geom_line(aes(x=dominant,y=de_sq_or, linetype = as.factor(fraction_blue))) + ggtitle("mincons"))

minlib <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_cl_or, colour = vl_cl_or),shape = "circle")  + geom_line(aes(x=dominant,y=dv_cl_or, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_cl_or, colour = et_cl_or),shape="square") + geom_line(aes(x=dominant,y=de_cl_or, linetype = as.factor(fraction_blue)))+ ggtitle("minlib"))

majcons
majlib
mincons
minlib

```

```{r sa_asym, fig.cap="Fig. 8 orig", echo = FALSE}

sa_asym <- read.csv("df_sa_asym.csv",sep = ",")

excl <- names(sa_asym) %in% c("run", "density","step")
df <- sa_asym[!excl] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

df <- df %>% filter(eth_con_min==20 & eth_con_loc==20 & val_lib_min==20)
df <- df %>% filter(fraction_blue==50|fraction_blue==60|fraction_blue==80)
df <- df %>% filter(circle_orange==50)

dominant <- df$val_lib_loc

majcons <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_sq_bl, colour = vl_sq_bl),shape = "circle")  + geom_line(aes(x=dominant,y=dv_sq_bl, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_sq_bl, colour = et_sq_bl),shape="square") + geom_line(aes(x=dominant,y=de_sq_bl, linetype = as.factor(fraction_blue)))+ ggtitle("majcons"))

majlib <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_cl_bl, colour = vl_cl_bl),shape = "circle")  + geom_line(aes(x=dominant,y=dv_cl_bl, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_cl_bl, colour = et_cl_bl),shape="square") + geom_line(aes(x=dominant,y=de_cl_bl, linetype = as.factor(fraction_blue))) + ggtitle("majlib"))

mincons <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_sq_or, colour = vl_sq_or),shape = "circle")  + geom_line(aes(x=dominant,y=dv_sq_or, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_sq_or, colour = et_sq_or),shape="square") + geom_line(aes(x=dominant,y=de_sq_or, linetype = as.factor(fraction_blue))) + ggtitle("mincons"))

minlib <- ggplotly(ggplot(data = df) + geom_point(aes(x=dominant,y=dv_cl_or, colour = vl_cl_or),shape = "circle")  + geom_line(aes(x=dominant,y=dv_cl_or, linetype = as.factor(fraction_blue))) +
           geom_point(aes(x=dominant,y=de_cl_or, colour = et_cl_or),shape="square") + geom_line(aes(x=dominant,y=de_cl_or, linetype = as.factor(fraction_blue)))+ ggtitle("minlib"))

majcons
majlib
mincons
minlib

```

# Comparing beta majority over beta minority liberal

```{r by-product_cons_original, fig.cap="Effect of liberals majority vs minority on value segregation conservatives", echo = FALSE}

# ethnic and value segregation for each group in the annex

##### 
# By-product value of conservatives by value preference liberals

# Majority

mech_cons <- mech %>% filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_min==0 | val_lib_min==1 |val_lib_min==7) %>% gather(key=ethnics, value=Clust, dv_sq_bl,dv_sq_or)

cons_maj <- ggplot(data = mech_cons, aes(x=val_lib_loc,y=Clust,color=  `ethnics`,linetype= as.factor(fraction_blue))) + geom_point(size=1.5, shape = "square") + geom_line() + 
  facet_wrap(~val_lib_min,dir="v",labeller = labeller(val_lib_min = valueminority.labs)) +
    scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  xlab("ß value liberal majority") + ylab("Spatial clustering") +
  geom_hline(yintercept=1,color="red") +
  scale_color_manual(values=c("blue","darkorange"),labels=c("Majority","Minority"),name="Ethnicity") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
   guides(color  = guide_legend(order = 1), linetype = guide_legend(order = 2)) +
  theme_bw()

# Minority

mech_lib <- mech %>%  filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_loc==0 | val_lib_loc==1 |val_lib_loc==7) %>% gather(key=ethnics, value=Cluste, dv_sq_bl,dv_sq_or)

cons_min <- ggplot(data = mech_lib, aes(x=val_lib_min,y=Cluste,color=  `ethnics`,linetype= as.factor(fraction_blue))) + geom_point(size=1.5, shape = "square") + geom_line() + 
  facet_wrap(~val_lib_loc,dir="v",labeller = labeller(val_lib_loc = valuemajority.labs)) +
  xlab("ß value liberal minority") + ylab("Spatial clustering") +
  scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  geom_hline(yintercept=1,color="red") +
  scale_color_manual(values=c("blue","darkorange"),labels=c("Majority","Minority"),name="Ethnicity") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
   guides(color  = guide_legend(order = 1), linetype = guide_legend(order = 2)) +
 theme_bw()

ggplotly(cons_maj)
ggplotly(cons_min)

lib_cons <- ggpubr::ggarrange(cons_maj,cons_min, common.legend = TRUE, legend = "bottom")

ggsave("lib_cons.jpg",plot=lib_cons)

```

```{r by-product ethnic cons, fig.cap="Effect of liberals majority vs minority on ethnic segregation conservatives", echo = FALSE}
# Not interesting, goes in the annex

# ethnic and value segregation for each group in the annex

##### 
# By-product value of conservatives by value preference liberals

# Majority

mech_cons <- mech %>% filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_min==0 | val_lib_min==1 |val_lib_min==7) %>% gather(key=ethnics, value=Clust, de_sq_bl,de_sq_or)

cons_maj_et <- ggplot(data = mech_cons, aes(x=val_lib_loc,y=Clust,color=  `ethnics`,linetype= as.factor(fraction_blue))) + geom_point(size=1.5, shape = "square") + geom_line() + 
  facet_wrap(~val_lib_min,dir="v",labeller = labeller(val_lib_min = valueminority.labs)) +
    scale_y_continuous(breaks= seq(0,5,by=0.5)) +
  xlab("ß value liberal majority") + ylab("Spatial clustering") +
  geom_hline(yintercept=1,color="red") +
  scale_color_manual(values=c("blue","darkorange"),labels=c("Majority","Minority"),name="Ethnicity") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
   guides(color  = guide_legend(order = 1), linetype = guide_legend(order = 2)) +
  theme_bw()

# Minority

mech_lib <- mech %>%  filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_loc==0 | val_lib_loc==1 |val_lib_loc==7) %>% gather(key=ethnics, value=Cluste, de_sq_bl,de_sq_or)

cons_min_et <- ggplot(data = mech_lib, aes(x=val_lib_min,y=Cluste,color=  `ethnics`,linetype= as.factor(fraction_blue))) + geom_point(size=1.5, shape = "square") + geom_line() + 
  facet_wrap(~val_lib_loc,dir="v",labeller = labeller(val_lib_loc = valuemajority.labs)) +
  xlab("ß value liberal minority") + ylab("Spatial clustering") +
  scale_y_continuous(breaks= seq(0,5,by=0.5)) +
  geom_hline(yintercept=1,color="red") +
  scale_color_manual(values=c("blue","darkorange"),labels=c("Majority","Minority"),name="Ethnicity") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
   guides(color  = guide_legend(order = 1), linetype = guide_legend(order = 2)) +
 theme_bw()

ggplotly(cons_maj_et)
ggplotly(cons_min_et)

lib_cons_et <- ggpubr::ggarrange(cons_maj_et,cons_min_et, common.legend = TRUE, legend = "bottom")

#ggsave("lib_cons.jpg",plot=lib_cons)

```

```{r beta_maj, fig.cap="effect liberals one  dataset", echo = FALSE}

#####
# Ethnic and value of liberals above each other
mech_vlmj <- mech %>% filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_min==0 | val_lib_min==1 |val_lib_min==7) %>% gather(key = segregation, value = Clustmj, de_cl_bl, dv_cl_bl)

librmj <- ggplot(data=mech_vlmj,aes(x=val_lib_loc,y = Clustmj,shape=`segregation`,linetype=as.factor(fraction_blue))) + 
  geom_point(size=1.5,color="blue") + geom_line(color="blue") +
facet_wrap(~val_lib_min,dir="v",labeller=labeller(val_lib_min=valueminority.labs)) +
  xlab("ß value liberals majority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Ethnic","Value"),name="Segregation" ) +
 ggtitle("Liberals Majority") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

mech_vlmn <- mech %>%  filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(val_lib_loc==0 | val_lib_loc==1 |val_lib_loc==7) %>% gather(key = segregation, value = Clustmn, de_cl_or, dv_cl_or)

librmn <- ggplot(data=mech_vlmn,aes(x=val_lib_min,y = Clustmn,shape=`segregation`,linetype=as.factor(fraction_blue))) + 
  geom_point(size=1.5,color="orange") + geom_line(color="orange") +
facet_wrap(~val_lib_loc,dir="v",labeller=labeller(val_lib_loc=valuemajority.labs)) +
 scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  xlab("ß value liberals minority") + ylab("Spatial Clustering") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
  scale_shape_manual(values=c("square","circle"),labels=c("Ethnic","Value"),name="Segregation") +
    geom_hline(yintercept=1,color="red") +
    guides(  linetype = guide_legend(order = 1), shape = guide_legend(order = 2)) +
 #  guides(color=FALSE) + 
    ggtitle("Liberals Minority") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(librmj)
ggplotly(librmn)

lib_lib <- ggpubr::ggarrange(librmj,librmn, common.legend = TRUE, legend = "bottom")

ggsave("lib_lib.jpg",plot=lib_lib)

```

```{r etlib_min, fig.cap="Ethnic preference majority or minority over liberals or conservatives minority", echo = FALSE}

excl <- names(df_etlib) %in% c("run", "density","step")
df_etlib <- df_etlib[!excl] %>% group_by(ID,eth_con_loc,eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

## Majority

df_ethloc <- df_etlib %>% filter(ID == "etlib_loc") %>% filter(fraction_blue == 80) %>% gather(key = subgroup, value = Clustetmj, de_sq_or,dv_sq_or,de_cl_or,dv_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_or","de_cl_or"), "Value" = c("dv_sq_or","dv_cl_or") ),"Ethnic"),
              `orientation` = fct_relevel(fct_collapse(subgroup,"Conservative" = c("de_sq_or","dv_sq_or"), "Liberal" = c("de_cl_or","dv_cl_or")), "Conservative"))

maj_min <- ggplot(data=df_ethloc,aes(x=eth_lib_loc,y = Clustetmj,linetype=`segregation`,shape = `orientation`)) + 
  geom_point(size=1.5,color="orange") + geom_line(color="orange") +
facet_wrap(~eth_lib_min,dir="v",labeller=labeller(eth_lib_min=ethnicminority.labs)) +
  xlab("ß ethnic liberals majority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,5,by=0.50)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("Ethnic","Value"),name="Segregation" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Conservative","Liberal"),name="Orientation \n minority" ) +
 ggtitle("") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(maj_min)
   maj_min
## Minority

df_ethmin <- df_etlib %>% filter(ID == "etlib_min") %>% filter(fraction_blue == 80) %>% gather(key = subgroup, value = Clustetmn, de_sq_or,dv_sq_or,de_cl_or,dv_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_or","de_cl_or"), "Value" = c("dv_sq_or","dv_cl_or") ),"Ethnic"),
              `orientation` = fct_relevel(fct_collapse(subgroup,"Conservative" = c("de_sq_or","dv_sq_or"), "Liberal" = c("de_cl_or","dv_cl_or")), "Conservative"))

min_min <- ggplot(data=df_ethmin,aes(x=eth_lib_min,y = Clustetmn,linetype=`segregation`,shape = `orientation`)) + 
  geom_point(size=1.5,color="orange") + geom_line(color="orange") +
facet_wrap(~eth_lib_loc,dir="v",labeller=labeller(eth_lib_loc=ethnicmajority.labs)) +
  xlab("ß ethnic liberals minority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,5,by=0.50)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("Ethnic","Value"),name="Segregation" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Conservative","Liberal"),name="Orientation \n minority" ) +
 ggtitle("") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(min_min)
   min_min

eth_lib_min <- ggpubr::ggarrange(maj_min,min_min, common.legend = TRUE, legend = "bottom")

ggsave("eth_lib_min.jpg",plot=eth_lib_min)

```

```{r etlib_min_test fraction, fig.cap="Ethnic preference majority or minority over liberals or conservatives minority", echo = FALSE}

df_etlib <- read.csv("df_etlib.csv",sep=",")

excl <- names(df_etlib) %in% c("run", "density","step")
df_etlib <- df_etlib[!excl] %>% group_by(ID,eth_con_loc,eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

## Majority

df_ethloc <- df_etlib %>% filter(ID == "etlib_loc") %>% filter(fraction_blue == 50 | fraction_blue == 80) %>% gather(key = subgroup, value = Clustetmj, de_sq_or,dv_sq_or,de_cl_or,dv_cl_or) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_or","de_cl_or"), "Value" = c("dv_sq_or","dv_cl_or") ),"Ethnic"),
              `orientation` = fct_relevel(fct_collapse(subgroup,"Conservative" = c("de_sq_or","dv_sq_or"), "Liberal" = c("de_cl_or","dv_cl_or")), "Conservative"))

df_ethloc <- df_etlib %>% filter(ID == "etlib_loc") %>% filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(eth_lib_min==0 | eth_lib_min==1 |eth_lib_min==7) %>% gather(key = segregation, value = Clustmj, de_cl_bl, dv_cl_bl)

librmj <- ggplot(data=df_ethloc,aes(x=eth_lib_loc,y = Clustmj,shape=`segregation`,linetype=as.factor(fraction_blue))) + 
  geom_point(size=1.5,color="blue") + geom_line(color="blue") +
facet_wrap(~eth_lib_min,dir="v",labeller=labeller(eth_lib_min=ethnicminority.labs)) +
  xlab("ß ethnic liberals majority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Ethnic","Value"),name="Segregation" ) +
 ggtitle("Liberals Majority") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

df_etmin <- df_etlib %>% filter(ID == "etlib_min") %>% filter(fraction_blue == 50 | fraction_blue == 80) %>%  filter(eth_lib_loc==0 | eth_lib_loc==1 |eth_lib_loc==7) %>% gather(key = segregation, value = Clustmn, de_cl_or, dv_cl_or)

librmn <- ggplot(data=df_etmin,aes(x=eth_lib_min,y = Clustmn,shape=`segregation`,linetype=as.factor(fraction_blue))) + 
  geom_point(size=1.5,color="orange") + geom_line(color="orange") +
facet_wrap(~eth_lib_loc,dir="v",labeller=labeller(eth_lib_loc=ethnicmajority.labs)) +
 scale_y_continuous(breaks= seq(0,2,by=0.25)) +
  xlab("ß ethnic liberals minority") + ylab("Spatial Clustering") +
    scale_linetype_manual(values=c("solid","dashed"),labels=c("50%","80%"),name="% Majority") +
  scale_shape_manual(values=c("square","circle"),labels=c("Ethnic","Value"),name="Segregation") +
    geom_hline(yintercept=1,color="red") +
    guides(  linetype = guide_legend(order = 1), shape = guide_legend(order = 2)) +
 #  guides(color=FALSE) + 
    ggtitle("Liberals Minority") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(librmj)
ggplotly(librmn)

lib_lib_ethnic <- ggpubr::ggarrange(librmj,librmn, common.legend = TRUE, legend = "bottom")

# ggsave("lib_lib.jpg",plot=lib_lib)

```

```{r etlib_maj, fig.cap="Ethnic preference majority or minority over liberals or conservatives majority", echo = FALSE}

## Majority

df_ethloc_maj <- df_etlib %>% filter(ID == "etlib_loc") %>% filter(fraction_blue == 80) %>% gather(key = subgroup, value = Clustetmj, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl"), "Value" = c("dv_sq_bl","dv_cl_bl") ),"Ethnic"),
              `orientation` = fct_relevel(fct_collapse(subgroup,"Conservative" = c("de_sq_bl","dv_sq_bl"), "Liberal" = c("de_cl_bl","dv_cl_bl")), "Conservative"))

maj_maj <- ggplot(data=df_ethloc_maj,aes(x=eth_lib_loc,y = Clustetmj,linetype=`segregation`,shape = `orientation`)) + 
  geom_point(size=1.5,color="blue") + geom_line(color="blue") +
facet_wrap(~eth_lib_min,dir="v",labeller=labeller(eth_lib_min=ethnicminority.labs)) +
  xlab("ß ethnic liberals majority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,5,by=0.50)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("Ethnic","Value"),name="Segregation" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Conservative","Liberal"),name="Orientation \n majority" ) +
 ggtitle("") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(maj_maj)
   maj_maj
## Minority

df_ethmin_maj <- df_etlib %>% filter(ID == "etlib_min") %>% filter(fraction_blue == 80) %>% gather(key = subgroup, value = Clustetmn, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl) %>% mutate(`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl"), "Value" = c("dv_sq_bl","dv_cl_bl") ),"Ethnic"),
              `orientation` = fct_relevel(fct_collapse(subgroup,"Conservative" = c("de_sq_bl","dv_sq_bl"), "Liberal" = c("de_cl_bl","dv_cl_bl")), "Conservative"))

min_maj <- ggplot(data=df_ethmin_maj,aes(x=eth_lib_min,y = Clustetmn,linetype=`segregation`,shape = `orientation`)) + 
  geom_point(size=1.5,color="blue") + geom_line(color="blue") +
facet_wrap(~eth_lib_loc,dir="v",labeller=labeller(eth_lib_loc=ethnicmajority.labs)) +
  xlab("ß ethnic liberals minority") + ylab("Spatial Clustering") +
  scale_y_continuous(breaks= seq(0,5,by=0.50)) +
  scale_linetype_manual(values=c("solid","dashed"),labels=c("Ethnic","Value"),name="Segregation" ) +
  scale_shape_manual(values=c("square","circle"),labels=c("Conservative","Liberal"),name="Orientation \n majority" ) +
 ggtitle("") +
    geom_hline(yintercept=1,color="red") +
      theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(min_maj)
   min_maj

eth_lib_maj <- ggpubr::ggarrange(maj_maj,min_maj, common.legend = TRUE, legend = "bottom")

ggsave("eth_lib_maj.jpg",plot=eth_lib_maj)

```

```{r sec_not_rel, fig.cap="Value conservativess and ethnic liberal not primary", echo = FALSE}

df <- df_sec  %>%  gather(key= subgroup, value = Prox, de_sq_bl, dv_sq_bl, de_cl_bl, dv_cl_bl, de_sq_or, dv_sq_or, de_cl_or, dv_cl_or) %>% mutate(  `segregation` = fct_relevel(fct_collapse(subgroup,"ethnics" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "values" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"ethnics"),
 `orientation` = fct_relevel(fct_collapse(subgroup, "conservative" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "liberal" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")),"conservative" ), 
 `ethnia` = fct_relevel(fct_collapse(subgroup, "local" = c("de_sq_bl","dv_sq_bl","de_cl_bl","dv_cl_bl"),"minority" = c("de_sq_or","dv_sq_or","de_cl_or","dv_cl_or")), "local"))

dflow <- df %>% filter(ID == "valuecon_min" & fraction_blue == 80) %>% filter(RND == "low")
dfh <- df %>% filter(ID == "valuecon_min" & fraction_blue == 80) %>% filter(RND == "high")

low <- ggplot(data=dflow, aes(x=dflow$val_con_min,y=Prox, linetype = `segregation` , fill = `ethnia`,  shape = `orientation`))  +  geom_point(size= 1.5) +
  geom_line(aes(color = `ethnia`)) + 
  facet_wrap(~  dflow$val_con_loc) +
  ggtitle("Low randomness") + 
   scale_linetype_manual(values=c("solid","dashed"), labels = c("ethnic","value"), guide = guide_legend(title = "Segregation", order = 1)) +
  scale_fill_manual(values = c("blue","darkorange")) + 
 scale_color_manual(values = c("blue","darkorange"), guide = guide_legend(title="Ethnicity", order = 2)) + 
  scale_shape_manual(values = c(22,21), guide = guide_legend(title="Value" ,override.aes = list(size = 3), order = 3)) +
  guides(fill  = FALSE)

high <- ggplot(data=dfh, aes(x=dfh$val_con_min,y=Prox, linetype = `segregation` , fill = `ethnia`,  shape = `orientation`))  +  geom_point(size= 1.5) +
  geom_line(aes(color = `ethnia`)) + 
  facet_wrap(~  dfh$val_con_loc) +
  ggtitle("High randomness") + 
   scale_linetype_manual(values=c("solid","dashed"), labels = c("ethnic","value"), guide = guide_legend(title = "Segregation", order = 1)) +
  scale_fill_manual(values = c("blue","darkorange")) + 
 scale_color_manual(values = c("blue","darkorange"), guide = guide_legend(title="Ethnicity", order = 2)) + 
  scale_shape_manual(values = c(22,21), guide = guide_legend(title="Value" ,override.aes = list(size = 3), order = 3)) +
  guides(fill  = FALSE)

cmp <- ggpubr::ggarrange(low,high,common.legend = T)

```

```{r sec_test, fig.cap="Value conservativess and ethnic liberal not primary", echo = FALSE}

# pref <- names(df_sec) %in% c("eth_con_loc", "eth_con_min","val_loc_lib")

etlblc_high <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "etlib_loc", RND == "high", eth_lib_loc == 5 | eth_lib_loc == 0) %>% filter(eth_lib_min == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

etlblc_high_pic <- ggplot(etlblc_high, aes(fill = as.factor(etlblc_high$eth_lib_loc), y= etlblc_high$Clus, x = etlblc_high$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

etlblc_low <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "etlib_loc", RND == "low", eth_lib_loc == 20 | eth_lib_loc == 0) %>% filter(eth_lib_min == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

etlblc_low_pic <- ggplot(etlblc_low, aes(fill = as.factor(etlblc_low$eth_lib_loc), y= etlblc_low$Clus, x = etlblc_low$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

etlbmn_high <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "etlib_min", RND == "high", eth_lib_min == 5 | eth_lib_min == 0) %>% filter(eth_lib_loc == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

etlbmn_high_pic <- ggplot(etlbmn_high, aes(fill = as.factor(etlbmn_high$eth_lib_min), y= etlbmn_high$Clus, x = etlbmn_high$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

etlbmn_low <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "etlib_min", RND == "low", eth_lib_min == 20 | eth_lib_min == 0) %>% filter(eth_lib_loc == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

etlbmn_low_pic <- ggplot(etlbmn_low, aes(fill = as.factor(etlbmn_low$eth_lib_min), y= etlbmn_low$Clus, x = etlbmn_low$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

vlcnlc_low <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "valuecon_loc", RND == "low", val_con_loc == 20 | val_con_loc == 0) %>% filter(val_con_min == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

vlcnlc_low_pic <- ggplot(vlcnlc_low, aes(fill = as.factor(vlcnlc_low$val_con_loc), y= vlcnlc_low$Clus, x = vlcnlc_low$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

vlcnlc_high <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "valuecon_loc", RND == "high", val_con_loc == 5 | val_con_loc == 0) %>% filter(val_con_min == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

vlcnlc_high_pic <- ggplot(vlcnlc_high, aes(fill = as.factor(vlcnlc_high$val_con_loc), y= vlcnlc_high$Clus, x = vlcnlc_high$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

vlcnmn_low <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "valuecon_min", RND == "low", val_con_min == 20 | val_con_min == 0) %>% filter(val_con_loc == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

vlcnmn_low_pic <- ggplot(vlcnmn_low, aes(fill = as.factor(vlcnmn_low$val_con_min), y= vlcnmn_low$Clus, x = vlcnmn_low$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

vlcnmn_high <- df_sec  %>% filter(fraction_blue == 80) %>% filter(ID == "valuecon_min", RND == "high", val_con_min == 5 | val_con_min == 0) %>% filter(val_con_loc == 0) %>% gather(key = subgroup, value = Clus, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)

vlcnmn_high_pic <- ggplot(vlcnmn_high, aes(fill = as.factor(vlcnmn_high$val_con_min), y= vlcnmn_high$Clus, x = vlcnmn_high$subgroup) ) + 
    geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_y_continuous(limits= c(0,6),breaks=seq(0,6,by=0.5))

ggplotly(etlblc_high_pic)
ggplotly(etlblc_low_pic)
ggplotly(etlbmn_high_pic)
ggplotly(etlbmn_low_pic)
ggplotly(vlcnlc_low_pic)
ggplotly(vlcnlc_high_pic)
ggplotly(vlcnmn_low_pic)
ggplotly(vlcnmn_high_pic)

```

```{r 3D, fig.cap="3D pictures", echo = FALSE}

excl3D <- names(df_fin) %in% c("run", "density","step")
df_fin <- df_fin[!excl3D] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

vars <- c("de_sq_bl","dv_sq_bl","et_sq_bl","vl_sq_bl")
df_sqbl <- cbind(df_fin[,1:11],df_fin[,vars]) %>% filter(val_lib_loc==5,val_lib_min==5,eth_con_loc == 5,eth_con_min == 5)
df_high <- cbind(df_fin[,1:11],df_fin[,vars]) %>% filter(val_lib_loc==20,val_lib_min==20,eth_con_loc == 20,eth_con_min == 20)

lowE <- plot_ly(df_sqbl)  %>%
  add_trace(x=df_sqbl$fraction_blue,y=df_sqbl$circle_orange,z=df_sqbl$de_sq_bl,
            scene = "scene1",
                                        name = "Ethnic",
                                       type="scatter3d",
                                       mode="markers",
                                     marker = list(cmin=0,cmax=1,
                                                   # colorscale = "Reds",
                                     colorscale =  list(
                                        c(0,'pink'),
                                        c(0.25,'red'),
                                        c(0.5,'green'),
                                        c(0.75,'blue'),
                                        c(1, 'purple')),
                                       color= df_sqbl$et_sq_bl, showscale = T)

            )  %>% 
  layout(showlegend = F,
         title="Ethnic",
         scene = list(
      xaxis = list(title = "Ethnic Ratio",range=seq(50,90,by=10),dtick=10),
      yaxis = list(title = "% Liberal Minority",autorange = "reversed"),
      zaxis = list(title = "Cluster")))

 lowV<- plot_ly(df_sqbl)  %>%   add_trace(x=df_sqbl$fraction_blue,y=df_sqbl$circle_orange,z=df_sqbl$dv_sq_bl,
                                          scene =  "scene2",
                                     name = "Value",
                                     type="scatter3d",
                                     mode="markers",
                                    marker = list(cmin=0,cmax=1,
                                                   # colorscale = "Reds",
                                     colorscale =  list(
                                        c(0,'pink'),
                                        c(0.25,'red'),
                                        c(0.5,'green'),
                                        c(0.75,'blue'),
                                        c(1, 'purple')),
                                       color= df_sqbl$et_sq_bl, showscale = T)

            )  %>% 
       layout(showlegend = T,
         title="",
         scene = list(
      xaxis = list(title = "Ethnic Ratio",range=seq(50,90,by=10),dtick=10),
      yaxis = list(title = "% Liberal Minority",autorange = "reversed"),
      zaxis = list(title = "Cluster")))

A <- subplot(lowE,lowV) %>% layout(annotations = list(
 list(x = 0.2 , y = 1 , text = "Ethnic", showarrow = F, xref='paper', yref='paper'),
  list(x = 0.8 , y = 1, text = "Value", showarrow = F, xref='paper', yref='paper'))
)

   ###

high <- plot_ly(df_high)  %>% add_trace(x=df_high$fraction_blue,y=df_high$circle_orange,z=df_high$de_sq_bl,color= df_high$et_sq_bl,
                                       type="scatter3d",
                                       mode="markers",
                                       marker = list( colorscale = 'Reds', showscale = T))  %>%
  add_trace(x=df_high$fraction_blue,y=df_high$circle_orange,z=df_high$dv_sq_bl,color= df_high$vl_sq_bl,
              type="scatter3d",
                                       mode="markers",
                                      marker = list( colorscale = 'Greens', showscale = TRUE))  %>%
  layout(legend = list(orientation = 'h'),
         title="high",
         scene = list(
      xaxis = list(title = "Ethnic Ratio",range=seq(50,90,by=10),dtick=10),
      yaxis = list(title = "% Liberal Minority",autorange = "reversed"),
      zaxis = list(title = "Cluster")))

low
high

```

```{r 3Dsimple, fig.cap="3D pictures", echo = FALSE}

excl3D <- names(df_fin) %in% c("run", "density","step")
df_fin <- df_fin[!excl3D] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

vars <- c("de_sq_or","dv_sq_or","et_sq_or","vl_sq_or")
df_sqbl <- cbind(df_fin[,1:11],df_fin[,vars]) %>% filter(val_lib_loc==5,val_lib_min==5,eth_con_loc == 5,eth_con_min == 5)
df_high <- cbind(df_fin[,1:11],df_fin[,vars]) %>% filter(val_lib_loc==20,val_lib_min==20,eth_con_loc == 20,eth_con_min == 20)

low <- plot3D::scatter3D(x = df_sqbl$fraction_blue, y = df_sqbl$circle_orange, z = df_sqbl$de_sq_or, colvar = df_sqbl$et_sq_or, col = ramp.col (col = c("grey", "blue"), n = 100, alpha = 1), pch=20, cex= 2, phi = 0)

```

### Back

```{r to write 3D csv, delete}
# trd <- read.csv("vlmaj_ethrat_3D_1.csv",sep=",",skip = 6)
# trd2 <- read.csv("vlmaj_ethrat_3D_2.csv",sep=",",skip = 6)
# lgnd <- read.csv("legend_sa_asym.csv",sep=";")[,1]
# names(trd2) <- lgnd
# 
# trd_f <- rbind(trd,trd2)
# 
# trdmin <- read.csv("vlmin_ethrat_3D_1.csv",sep=",",skip = 6)
# trd2min <- read.csv("vlmin_ethrat_3D_2.csv",sep=",",skip = 6)
# names(trdmin) <- lgnd
# names(trd2min) <- lgnd
# 
# trdmin_f <- rbind(trdmin,trd2min)
# 
# etlib_mj <- read.csv("etlib_loc_3D_1.csv",sep=",",skip = 6)
# lget <- read.csv("legend_etlib.csv",sep=";")[,1]
# names(etlib_mj) <- lget
# 
# 
# order <- c("run","density","fraction_blue", "circle_blue","circle_orange" ,"eth_con_loc","val_lib_loc", "eth_con_min" , "val_lib_min" ,  "val_con_loc","eth_lib_loc","val_con_min" ,  "eth_lib_min",   "step"  , "et_sq_bl" , "et_cl_bl" ,"et_sq_or","et_cl_or","vl_sq_bl","vl_cl_bl","vl_sq_or","vl_cl_or","de_sq_bl","de_cl_bl","de_sq_or","de_cl_or","dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")
# 
# etlib_mj <- etlib_mj[,order]
# 
# etlib_mj2 <- read.csv("etlib_loc_3D_2.csv",sep=",",skip = 6)
# names(etlib_mj2) <- lget
# 
# etlib_mj2 <- etlib_mj2[,order]
# 
# df <- rbind(trdmin_f,etlib_mj,etlib_mj2)
# 
# etmin <- read.csv("etlib_min_3D_1.csv",sep = ",",skip = 6)
# names(etmin) <- lget
# 
# etmin2 <- read.csv("etlib_min_3D_2.csv",sep = ",",skip = 6)
# names(etmin2) <- lget
# 
# etmin_all <- rbind(etmin,etmin2)
# 
# etmin_all <- etmin_all[,order]
# 
# df_f <- rbind(df,etmin_all)
# 
# write.csv(df_f,"3D.csv",row.names = F)
```

```{r beta_maj, fig.cap="x-axis value liberals maj vs min, each group", echo = FALSE}
#####
# effect on ethnic and value preference of conservatives

consmj_majlib <- ggplot(data=mech_cons) + geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_sq_bl,color=mech_cons$et_sq_bl),shape=15)  + 
  geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_sq_bl,linetype=as.factor(mech_cons$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_bl,color=mech_cons$vl_sq_bl),shape =15) +
    geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_bl,linetype=as.factor(mech_cons$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_cons$val_lib_min,dir="v") +
  theme(legend.position = "bottom")

consmj_minlib <- ggplot(data=mech_lib) + geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$de_sq_bl,color=mech_lib$et_sq_bl),shape=15)  + 
  geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$de_sq_bl,linetype=as.factor(mech_lib$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_bl,color=mech_lib$vl_sq_bl),shape =15) +
    geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_bl,linetype=as.factor(mech_lib$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_lib$val_lib_loc,dir="v") +
  theme(legend.position = "bottom")

ggplotly(consmj_majlib)
ggplotly(consmj_minlib)

ggpubr::ggarrange(consmj_majlib,consmj_minlib)

##

consmn_majlib <- ggplot(data=mech_cons) + geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_sq_or,color=mech_cons$et_sq_or),shape=15)  + 
  geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_sq_or,linetype=as.factor(mech_cons$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
  scale_y_continuous(limits=c(0.5,5),breaks = seq(0,5,by=1)) +
geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_or,color=mech_cons$vl_sq_or),shape =15) +
    geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_or,linetype=as.factor(mech_cons$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_cons$val_lib_min,dir="v") +
  theme(legend.position = "bottom")

consmn_minlib <- ggplot(data=mech_lib) + geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$de_sq_or,color=mech_lib$et_sq_or),shape=15)  + 
  geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$de_sq_or,linetype=as.factor(mech_lib$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
   scale_y_continuous(limits=c(0.9,5),breaks = seq(0,5,by=1)) +
geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_or,color=mech_lib$vl_sq_or),shape =15) +
    geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_or,linetype=as.factor(mech_lib$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_lib$val_lib_loc,dir="v") +
  theme(legend.position = "bottom")

ggplotly(consmn_majlib)
ggplotly(consmn_minlib)

ggpubr::ggarrange(consmn_majlib,consmn_minlib)

#####
#behavior liberals

libmj_maj <- ggplot(data=mech_cons) + geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_cl_bl,color=mech_cons$et_cl_bl),shape="square")  + geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_cl_bl,linetype=as.factor(mech_cons$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_cl_bl,color=mech_cons$vl_cl_bl),shape ="circle") +
    geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_cl_bl,linetype=as.factor(mech_cons$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_cons$val_lib_min,dir="v") +
  theme(legend.position = "bottom")

libmj_min <-  ggplot(data=mech_lib) + geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$de_cl_bl,color=mech_lib$et_cl_bl),shape="square")  + 
  geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$de_cl_bl,linetype=as.factor(mech_lib$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_cl_bl,color=mech_lib$vl_cl_bl),shape ="circle") +
    geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_cl_bl,linetype=as.factor(mech_lib$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_lib$val_lib_loc,dir="v") +
  theme(legend.position = "bottom")

ggplotly(libmj_maj)
ggplotly(libmj_min)

ggpubr::ggarrange(libmj_maj,libmj_min)

libmn_maj <- ggplot(data=mech_cons) + geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_cl_or,color=mech_cons$et_cl_or),shape="square")  + geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$de_cl_or,linetype=as.factor(mech_cons$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_cl_or,color=mech_cons$vl_cl_or),shape ="circle") +
    geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_cl_or,linetype=as.factor(mech_cons$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_cons$val_lib_min,dir="v") +
  theme(legend.position = "bottom")

libmn_min <-  ggplot(data=mech_lib) + geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$de_cl_or,color=mech_lib$et_cl_or),shape="square")  + 
  geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$de_cl_or,linetype=as.factor(mech_lib$fraction_blue)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_cl_or,color=mech_lib$vl_cl_or),shape ="circle") +
    geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_cl_or,linetype=as.factor(mech_lib$fraction_blue)),color="darkorange") +
  facet_wrap(~mech_lib$val_lib_loc,dir="v") +
  theme(legend.position = "bottom")

ggplotly(libmn_maj)
ggplotly(libmn_min)

ggpubr::ggarrange(libmj_maj,libmj_min)
ggpubr::ggarrange(libmn_maj,libmn_min)

#####
# too complicated

lib_loc <- ggplot(data=mech_cons) + geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_bl,color=mech_cons$vl_sq_bl),shape=15)  + 
  geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_bl,linetype=as.factor(mech_cons$val_lib_min)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_or,color=mech_cons$vl_sq_or),shape =15) +
    geom_line(aes(x=mech_cons$val_lib_loc,y=mech_cons$dv_sq_or,linetype=as.factor(mech_cons$val_lib_min)),color="darkorange") +
  facet_wrap(~mech_cons$fraction_blue,dir="v") +
  theme(legend.position = "bottom")

lib_min <- ggplot(data=mech_lib) + geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_bl,color=mech_lib$vl_sq_bl),shape=15)  + 
  geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_bl,linetype=as.factor(mech_lib$val_lib_loc)),color="blue") +
   scale_color_gradient(low = "grey", high = "brown") +
geom_point(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_or,color=mech_lib$vl_sq_or),shape =15) +
    geom_line(aes(x=mech_lib$val_lib_min,y=mech_lib$dv_sq_or,linetype=as.factor(mech_lib$val_lib_loc)),color="darkorange") +
  facet_wrap(~mech_lib$fraction_blue,dir="v") +
  theme(legend.position = "bottom")

ggplotly(lib_loc)
ggplotly(lib_min)
ggpubr::ggarrange(lib_loc,lib_min,common.legend = TRUE)

#####
# fraction: take 50 65 80, beta = 0, 1, 7

df_min <- read.csv("vlmin_ethrat_1.csv",sep = ",",skip = 6)
legend <- read.csv("legend_sa_asym.csv",sep = ";")
legend <- as.data.frame(legend[,-2])
names(df_min) <- pull(legend)

df_maj <- read.csv("vlmaj_ethrat_1.csv",sep = ",",skip = 6)
names(df_maj) <- pull(legend)

excl <- names(df_min) %in% c("run", "density","step")
df_min <- df_min[!excl] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

excl <- names(df_maj) %in% c("run", "density","step")
df_maj <- df_maj[!excl] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)

df_min <- df_min %>% filter(eth_con_min==20 & eth_con_loc==20)
# df_min <- df_min %>% filter(fraction_blue==50 |  fraction_blue==80)
df_min <- df_min %>% filter(val_lib_loc == 0 | val_lib_loc == 1 |  val_lib_loc == 7 )

# val_lib_loc == 3 ||val_lib_loc == 7 |

df_maj <- df_maj %>% filter(eth_con_min==20 & eth_con_loc==20)
# df_maj <- df_maj %>% filter(fraction_blue==50 |  fraction_blue==80)
df_maj <- df_maj %>% filter(val_lib_min == 0 | val_lib_min == 1 |   val_lib_min == 7 )

#  val_lib_min == 3 | val_lib_min == 7

majcons_min <- ggplot(data = mech_lib) + geom_point(aes(x=val_lib_min,y=dv_sq_bl, colour = vl_sq_bl),shape = "circle",size=2)  + geom_line(aes(x=val_lib_min,y=dv_sq_bl, linetype = as.factor(val_lib_loc))) + facet_wrap(~mech_lib$fraction_blue) +
  # geom_point(aes(x=val_lib_min,y=de_sq_bl, colour = et_sq_bl),shape = "square",size=2)  + geom_line(aes(x=val_lib_min,y=de_sq_bl, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue)  + 
  ggtitle("majcons over min")

majcons_maj <- ggplot(data = mech_cons) + geom_point(aes(x=val_lib_loc,y=dv_sq_bl, colour = vl_sq_bl),shape = "circle",size=2)  + geom_line(aes(x=val_lib_loc,y=dv_sq_bl, linetype = as.factor(val_lib_min))) + facet_wrap(~mech_cons$fraction_blue) +
 # geom_point(aes(x=val_lib_loc,y=de_sq_bl, colour = et_sq_bl),shape = "square",size=2)  + geom_line(aes(x=val_lib_loc,y=de_sq_bl, linetype = as.factor(val_lib_min))) + facet_wrap(~df_min$fraction_blue)  + 
  ggtitle("majcons over min")

mincons_min <- ggplot(data = df_min) + geom_point(aes(x=val_lib_min,y=dv_sq_or, colour = vl_sq_or),shape = "circle",size=2)  + geom_line(aes(x=val_lib_min,y=dv_sq_or, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue) +
#   geom_point(aes(x=val_lib_min,y=de_sq_or, colour = et_sq_or),shape = "square",size=2)  + geom_line(aes(x=val_lib_min,y=de_sq_or, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue) +
  ggtitle("mincons over min")

mincons_maj <- ggplot(data = df_maj) + geom_point(aes(x=val_lib_loc,y=dv_sq_or, colour = vl_sq_or),shape = "circle",size=2)  + geom_line(aes(x=val_lib_loc,y=dv_sq_or, linetype = as.factor(val_lib_min))) + facet_wrap(~df_maj$fraction_blue) + # geom_point(aes(x=val_lib_loc,y=de_sq_or, colour = et_sq_or),shape = "square",size=2)  + geom_line(aes(x=val_lib_loc,y=de_sq_or, linetype = as.factor(val_lib_min))) + facet_wrap(~df_maj$fraction_blue) +
  ggtitle("mincons over maj")

ggplotly(majcons_min)
ggplotly(majcons_maj)

ggplotly(mincons_min)
ggplotly(mincons_maj)

majlib_min <- ggplot(data = df_min) + geom_point(aes(x=val_lib_min,y=dv_cl_bl, colour = vl_cl_bl),shape = "circle",size=2)  + geom_line(aes(x=val_lib_min,y=dv_cl_bl, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue)+ 
  # geom_point(aes(x=val_lib_min,y=de_cl_bl, colour = et_cl_bl),shape = "square",size=2)  + geom_line(aes(x=val_lib_min,y=de_cl_bl, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue)+ 
 ggtitle("majlib over min")

majlib_maj <- ggplot(data = df_maj) + geom_point(aes(x=val_lib_loc,y=dv_cl_bl, colour = vl_cl_bl),shape = "circle",size=2)  + geom_line(aes(x=val_lib_loc,y=dv_cl_bl, linetype = as.factor(val_lib_min))) + facet_wrap(~df_maj$fraction_blue) +
  # geom_point(aes(x=val_lib_loc,y=de_cl_bl, colour = et_cl_bl),shape = "square",size=2)  + geom_line(aes(x=val_lib_loc,y=de_cl_bl, linetype = as.factor(val_lib_min))) + facet_wrap(~df_min$fraction_blue)+
 ggtitle("majlib over maj")

ggplotly(majlib_min)
ggplotly(majlib_maj)

minlib_min <- ggplot(data = df_min) + geom_point(aes(x=val_lib_min,y=dv_cl_or, colour = vl_cl_or),shape = "circle",size=2)  + geom_line(aes(x=val_lib_min,y=dv_cl_or, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue)+ geom_point(aes(x=val_lib_min,y=de_cl_or, colour = et_cl_or),shape = "square",size=2)  + geom_line(aes(x=val_lib_min,y=de_cl_or, linetype = as.factor(val_lib_loc))) + facet_wrap(~df_min$fraction_blue)  +
  ggtitle("minlib over min")

minlib_maj <- ggplot(data = df_maj) + geom_point(aes(x=val_lib_loc,y=dv_cl_or, colour = vl_cl_or),shape = "circle",size=2)  + geom_line(aes(x=val_lib_loc,y=dv_cl_or, linetype = as.factor(val_lib_min))) + facet_wrap(~df_maj$fraction_blue) + geom_point(aes(x=val_lib_loc,y=de_cl_or, colour = et_cl_or),shape = "square",size=2)  + geom_line(aes(x=val_lib_loc,y=de_cl_or, linetype = as.factor(val_lib_min))) + facet_wrap(~df_maj$fraction_blue) +
  ggtitle("minlib over maj")

ggplotly(minlib_min)
ggplotly(minlib_maj)

```

```{r back}

# fraction: take 50 65 80, beta = 0, 1, 7

# vlmin1 <- read.csv("vlmin_ethrat_1.csv",sep = ",",skip = 6)
# legend <- read.csv("legend_sa_asym.csv",sep = ";")
# legend <- as.data.frame(legend[,-2])
# names(vlmin1) <- pull(legend)
# 
# vlmin2 <- read.csv("vlmin_ethrat_2.csv",sep = ",",skip = 6)
# names(vlmin2) <- pull(legend)
# 
# vlmaj1 <- read.csv("vlmaj_ethrat_1.csv",sep = ",",skip = 6)
# names(vlmaj1) <- pull(legend)
# 
# vlmaj2 <- read.csv("vlmaj_ethrat_2.csv",sep = ",",skip = 6)
# names(vlmaj2) <- pull(legend)
# 
# mech_lib <- rbind(vlmin1,vlmin2,vlmaj1,vlmaj2)
# write.csv(mech_lib,"mech_lib.csv",row.names = FALSE)

df_long <- df %>% gather(key= subgroup, value = Outcome, de_sq_bl, dv_sq_bl, de_cl_bl, dv_cl_bl, de_sq_or, dv_sq_or, de_cl_or, dv_cl_or
                         ) %>% mutate(

   # `measureseg` = fct_relevel(fct_collapse(subgroup,"exposure" = c("et_sq_bl", "vl_sq_bl", "et_cl_bl", "vl_cl_bl", "et_sq_or", "vl_sq_or", "et_cl_or", "vl_cl_or"), "clustering" = c("de_sq_bl", "dv_sq_bl", "de_cl_bl", "dv_cl_bl", "de_sq_or", "dv_sq_or", "de_cl_or", "dv_cl_or")),"exposure"),

  `segregation` = fct_relevel(fct_collapse(subgroup,"ethnics" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or" ), "values" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or" )),"ethnics"),

 `orientation` = fct_relevel(fct_collapse(subgroup, "conservative" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "liberal" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")),"conservative" ),

 `ethnia` = fct_relevel(fct_collapse(subgroup, "local" = c("de_sq_bl","dv_sq_bl","de_cl_bl","dv_cl_bl"),"minority" = c("de_sq_or","dv_sq_or","de_cl_or","dv_cl_or")), "local"),

 `groupp` = fct_relevel(fct_collapse(subgroup, "cons_maj" = c("de_sq_bl","dv_sq_bl"),"lib_maj"=c("de_cl_bl","dv_cl_bl"),"cons_min"=c("de_sq_or","dv_sq_or"),"lib_min"=c("de_cl_or","dv_cl_or")),"cons_maj")
  )

## df_long 2 measures

df_long <- df %>% gather(key= subgroup, value = Outcome, et_sq_bl, vl_sq_bl, et_cl_bl, vl_cl_bl, et_sq_or, vl_sq_or, et_cl_or, vl_cl_or,
                          de_sq_bl, dv_sq_bl, de_cl_bl, dv_cl_bl, de_sq_or, dv_sq_or, de_cl_or, dv_cl_or ) %>% mutate(

   `measureseg` = fct_relevel(fct_collapse(subgroup,"exposure" = c("et_sq_bl", "vl_sq_bl", "et_cl_bl", "vl_cl_bl", "et_sq_or", "vl_sq_or", "et_cl_or", "vl_cl_or"), "clustering" = c("de_sq_bl", "dv_sq_bl", "de_cl_bl", "dv_cl_bl", "de_sq_or", "dv_sq_or", "de_cl_or", "dv_cl_or")),"exposure"),

  `segregation` = fct_relevel(fct_collapse(subgroup,"ethnics" = c("et_sq_bl","et_cl_bl","et_sq_or","et_cl_or","de_sq_bl","de_cl_bl","de_sq_or", "de_cl_or" ), "values" = c("vl_sq_bl","vl_cl_bl","vl_sq_or","vl_cl_or","dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or" )),"ethnics"),

 `orientation` = fct_relevel(fct_collapse(subgroup, "conservative" = c("et_sq_bl","vl_sq_bl","et_sq_or","vl_sq_or","de_sq_bl", "dv_sq_bl", "de_sq_or", "dv_sq_or"), "liberal" = c("et_cl_bl","vl_cl_bl","et_cl_or","vl_cl_or","de_cl_bl", "dv_cl_bl","de_cl_or", "dv_cl_or")),"conservative" ), 
 `ethnia` = fct_relevel(fct_collapse(subgroup, "local" = c("et_sq_bl","vl_sq_bl","et_cl_bl","vl_cl_bl","de_sq_bl", "dv_sq_bl", "de_cl_bl", "dv_cl_bl"),"minority" = c("et_sq_or","vl_sq_or","et_cl_or","vl_cl_or","de_sq_or", "dv_sq_or", "de_cl_or", "dv_cl_or")), "local")
 )

## 3D plotly

dfcl <- df_long[which(df_long$measureseg=="clustering"),]

df_etsqbl<- df[,which((names(df) %in% c("fraction_blue","dominant","et_sq_bl")))]

a <- as.matrix(df$dominant)
class(a)

b <- as.matrix(df$fraction_blue)
z <- as.matrix(df$et_sq_bl)

diba <- data.matrix(df)
class(diba)

M <- mesh(seq(0, 6*pi, length.out = 50),seq(pi/3, pi, length.out = 50))
u <- M$x ; v <- M$y

x <- data.matrix(df$fraction_blue)
y <- data.matrix(df$dominant)
z <- data.matrix(df$et_sq_bl)

dfa <- df[which(df$dominant <= 15),]
#plot_ly(data=df, x=df$dominant, y=df$fraction_blue, z=df$de_sq_bl, color = df$et_sq_bl) +
  plot_ly(data=dfa) %>% 
    add_trace(x=dfa$dominant, y=dfa$fraction_blue, z=dfa$de_sq_or, color = dfa$et_sq_bl,type = 'scatter3d',mode = 'lines+markers', marker = list( symbol="square") ) %>% 
     add_trace(x=dfa$dominant, y=dfa$fraction_blue, z=dfa$de_sq_bl, color = dfa$et_sq_bl, type = 'scatter3d',mode = 'lines+markers') %>%
    #   add_trace(x=df$dominant, y=df$fraction_blue, z=df$de_cl_or, color = df$et_cl_or, type = 'scatter3d',mode = 'lines+markers') %>%
    #     add_trace(x=df$dominant, y=df$fraction_blue, z=df$de_cl_bl, color = df$et_cl_bl, type = 'scatter3d',mode = 'lines+markers') %>%

    layout(
  #  title = "Layout options in a 3d scatter plot",
    scene = list(
      xaxis = list(title = "dominant"),
      yaxis = list(title = "ethnic ratio"),
      zaxis = list(title = "cluster")
    ))

# plot for each subgroup, aesthetic: x-axis:beta, y-axis:spatial clustering, linetype  = fraction_blue, shape = type of segregation, bright = exposure

## make ggplot into plotly
x1 <- 1:100
x2 <- x1 + 100
x3 <- x2 + 100

x <-  c(x1, x2, x3)
y <- c(2*x1, 5*x2, -2*x3)

group <-  c(rep("A", length(x1)),
           rep("B", length(x2)),
           rep("C", length(x3)))

ds <- data.frame(x, y, group)

# Use the group aesthetic to ensure lines are drawn separately for each group
p <- ggplot(ds, aes(x, y)) +
  geom_line(aes(group = group, color = group), size = 2) +
  ggtitle("Group specific line chart")

fig <- ggplotly(p)

## regression fit to 3D

# data(mtcars)
# 
# mtcars$shape <- cut(mtcars$mpg,
#                     breaks = c(0,18, 26, 100),
#                     labels = c("square", "circle", "diamond"))
# mtcars$color <- cut(mtcars$mpg,
#                     breaks = c(0,18, 26, 100),
#                     labels = c("red", "yellow", "green"))
# 
# plot_ly(type = "scatter", data = mtcars, x = rownames(mtcars), y = mtcars$mpg,
#         mode = "markers", symbol = ~I(shape), color = ~I(color))

x <- data.matrix(df$dominant)
y <- data.matrix(df$fraction_blue)
z <- data.matrix(df$et_sq_bl)

fit <- lm(z ~ as.integer(x) + as.integer(y))
# predict values on regular xy grid
grid.lines = 100
x.pred <- seq(min(x), max(x), length.out = grid.lines)
y.pred <- seq(min(y), max(y), length.out = grid.lines)
xy <- expand.grid( x = x.pred, y = y.pred)
z.pred <- matrix(predict(fit, newdata = xy),
                 nrow = grid.lines, ncol = grid.lines)
# fitted points for droplines to surface
fitpoints <- predict(fit)
# scatter plot with regression plane
scatter3D(x, y, z, pch = 18, cex = 2,
    theta = 90, phi = 0, ticktype = "detailed",
    xlab = "dominant", ylab = "maj/min", zlab = "de_sq_mj",
    surf = list(x = x.pred, y = y.pred, z = z.pred,
    facets = NA, fit = fitpoints), main = "df")

```