### 2021-07-13 af2.Rmd
one file df, abm, analysis

 ---
title: "additional3D"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/rocpa/OneDrive/Desktop/AF2/analysis")

library(dplyr)
library(ggplot2)
library(plotly)
library(scatterplot3d)
library(reshape)
library(reshape2)
library(tidyr)
library(forcats)
library(plot3D)
library(xlsx)

# To combine dataframes
# dfg <- function(x,id){
# lg <- read.csv("code_dv.csv", sep = ";")
# x <- rbind(read.csv(paste0(x,".csv"),sep = ",",skip = 6),read.csv(paste0(x,".csv"),sep = ",",skip = 6))
# names(x) <- lg[,1]
# x$ID <- id
# as.data.frame(x)
# }
# 
# bsl <- dfg("bsl_dom_sec","bsl")
# bsl_dom <- dfg("bsl_dom","bsl_dom")
# bsl_sec_cl <- dfg("bsl_dom_sec_circle","bsl_sec_cl")
# bsl_sec_sq <- dfg("bsl_dom_sec_square","bsl_sec_sq")
# sens_lib  <- dfg("sens_lib" ,"sens_lib")
# sens_con  <- dfg( "sens_con" ,"sens_con")
# bsl_lib_con   <- dfg("bsl_value_lib_con","bsl_lib_con")
# bsl_con_lib  <- dfg("bsl_value_con_lib","bsl_con_lib")
# hmp  <- dfg("bsl_dom_sec_htmp","hmp")
# asm  <- dfg("asym_bsl_sec","asm")
# asm_dom  <- dfg("asym_bsl_secno","asm_dom")
# libmj  <- dfg("asym_libmj_mn","libmj")
# libmn  <- dfg("asym_libmn_mj","libmn")
# dislib  <- dfg("asym_dislib","dislib")
# dislib_maj  <- dfg("asym_dislib_libmajL","dislib_maj")
# dislib_min  <- dfg("asym_dislib_libminL","dislib_min")
# etlib_bsl  <- dfg("asym_et_bsl","etlib_bsl")
# etlib_maj  <- dfg("asym_et_mj","etlib_maj")
# etlib_min  <- dfg("asym_et_mn","etlib_min")
# dislibet_maj  <- dfg("asym_dislib_et_mj", "dislibet_maj")
# dislibet_min  <- dfg("asym_dislib_et_mn","dislibet_min")
# 
# df_tot <- rbind(bsl,bsl_dom, bsl_sec_cl, bsl_sec_sq,sens_lib,sens_con,bsl_lib_con ,bsl_con_lib ,hmp,  asm , asm_dom , libmj , libmn  ,dislib , dislib_maj,dislib_min  ,etlib_bsl,etlib_maj ,etlib_min ,dislibet_maj ,dislibet_min)
# 
# 
# write.csv(df_tot,"df_tot.csv",row.names = F)

df <- read.csv("df_tot.csv",sep =",")

# 
# 
# 
# 
# df_3D <- read.csv("3D.csv",sep=",")
# excl3D <- names(df_3D) %in% c("run", "density","step")
# df_3D <- df_3D[!excl3D] %>% group_by(eth_con_loc, eth_con_min,eth_lib_loc,eth_lib_min,val_lib_loc,val_lib_min,val_con_loc,val_con_min,circle_orange,circle_blue,fraction_blue) %>% summarise_all(mean)
# 
# 
# value_basic_den <- read.csv("value_basic_den.csv",sep = ",")
# excl <- names(value_basic_den) %in% c("run","density","step")
# value_basic_den <- value_basic_den[!excl] %>% group_by(ID,fraction_blue,circle_blue,circle_orange,circle,square) %>% summarise_all(mean)
# 
# 
# value_dominant_den <- read.csv("value_dominant_den.csv",sep = ",")
# excl <- names(value_dominant_den) %in% c("run","density","step")
# value_dominant_den <- value_dominant_den[!excl] %>% group_by(ID,fraction_blue,circle_blue,circle_orange,dominant_square,dominant_circle,secondary_square,secondary_circle) %>% summarise_all(mean) 
#

```

```{r test_val-eth, fig.cap= "Checking effect of lvalue liberal = ethnic liberal" echo =  FALSE}

# equal size; liberals: value = ethnic

df <- read.csv("val_eth_circle3.csv",skip=6,sep=",")
lg <- read.csv("lg_val_eth_circle3.csv",sep =";")
names(df) <- lg[,1]

df <- df %>%  gather(key = subgroup, value = Segra, et_sq,vl_sq,et_cl,vl_cl,den_sq,den_cl)  %>% mutate(
`outcome` = fct_relevel(fct_collapse(subgroup,"Ethnic Segregation" = c("et_sq","et_cl"), "Value Segregation" = c("vl_sq","vl_cl"), "Density" = c("den_sq","den_cl")),"Ethnic Segregation"), 
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("et_sq","vl_sq","den_sq"), "Liberals" = c("et_cl","vl_cl","den_cl")), "Conservatives"))

df %>% filter(dominant_square == circle_both) %>% ggplot(aes(x= circle_both,y = Segra, color = `outcome`, shape = `orientation`)) + geom_point() + geom_line() +
  scale_color_manual(values = c("green","gray","purple"), guide = guide_legend(title = "Outcome")) +
  scale_shape_manual(values = c("square","circle"), guide = guide_legend(title = "Orientation"))  +
  scale_y_continuous(limits=c(0.45,1), breaks=seq(0.5,1,by=0.1)) +
  xlab("ß value liberals = ß ethnic liberals") +
  ylab("Outcome") +
  theme_bw()+
  theme(legend.position = "bottom")

```

# Basic two preferences
```{r dompref, echo = FALSE}

value_dominant_den <- value_dominant_den %>%  gather(key = subgroup, value = Segra, et_sq,vl_sq,et_cl,vl_cl,den_sq,den_cl)  %>% mutate(
`outcome` = fct_relevel(fct_collapse(subgroup,"Ethnic Segregation" = c("et_sq","et_cl"), "Value Segregation" = c("vl_sq","vl_cl"), "Density" = c("den_sq","den_cl")),"Ethnic Segregation"), 
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("et_sq","vl_sq","den_sq"), "Liberals" = c("et_cl","vl_cl","den_cl")), "Conservatives"))

value_basic_den <- value_basic_den %>%  gather(key = subgroup, value = Segra, et_sq,vl_sq,et_cl,vl_cl,den_sq,den_cl)  %>% mutate(
`outcome` = fct_relevel(fct_collapse(subgroup,"Ethnic Segregation" = c("et_sq","et_cl"), "Value Segregation" = c("vl_sq","vl_cl"), "Density" = c("den_sq","den_cl")),"Ethnic Segregation"), 
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("et_sq","vl_sq","den_sq"), "Liberals" = c("et_cl","vl_cl","den_cl")), "Conservatives"))

lib_dom <- value_dominant_den %>% filter (ID == "circle") %>% ggplot(aes(x = dominant_circle, y = Segra, color = `outcome`, shape = `orientation` )) + geom_point(size = 2) + geom_line() +
  scale_color_manual(values = c("green","gray","purple"), guide = guide_legend(title = "Outcome")) +
  scale_shape_manual(values = c("square","circle"), guide = guide_legend(title = "Orientation"))  +
  scale_y_continuous(limits=c(0.45,1), breaks=seq(0.5,1,by=0.1)) +
  xlab("ß value liberals \nall other ß = 0") +
  ylab("Outcome") +
  theme_bw()

lib_both <- value_basic_den %>% filter (ID == "circle") %>% ggplot(aes(x = circle, y = Segra, color = `outcome`, shape = `orientation` )) + geom_point(size = 2) + geom_line() +
  scale_color_manual(values = c("green","gray","purple"), guide = guide_legend(title = "Outcome")) +
  scale_shape_manual(values = c("square","circle"), guide = guide_legend(title = "Orientation"))  +
   scale_y_continuous(limits=c(0.45,1), breaks=seq(0.5,1,by=0.1)) +
  xlab("ß value liberals = ß ethnic liberals") +
  ylab("Outcome") +
  theme_bw()

cons_dom <- value_dominant_den %>% filter (ID == "square" & dominant_circle == 0) %>% ggplot(aes(x = dominant_square, y = Segra, color = `outcome`, shape = `orientation` )) + geom_point(size = 2) + geom_line() +
  scale_color_manual(values = c("green","gray","purple"), guide = guide_legend(title = "Outcome")) +
  scale_shape_manual(values = c("square","circle"), guide = guide_legend(title = "Orientation"))  +
     scale_y_continuous(limits=c(0.45,1), breaks=seq(0.5,1,by=0.1)) +
  xlab("ß ethnic conservatives \nall other ß = 0") +
  ylab("Outcome") +
  theme_bw()

cons_both <- value_basic_den %>% filter (ID == "square") %>% ggplot(aes(x = square, y = Segra, color = `outcome`, shape = `orientation` )) + geom_point(size = 2) + geom_line() +
  scale_color_manual(values = c("green","gray","purple"), guide = guide_legend(title = "Outcome")) +
  scale_shape_manual(values = c("square","circle"), guide = guide_legend(title = "Orientation"))  +
    scale_y_continuous(limits=c(0.45,1), breaks=seq(0.5,1,by=0.1)) +
  xlab("ß ethnic conservatives = ß value conservatives") +
  ylab("Outcome") +
  theme_bw()

a <- ggpubr::ggarrange(lib_dom,cons_dom , common.legend = TRUE, legend = "bottom")
b <- ggpubr::ggarrange(lib_both,cons_both, common.legend = TRUE, legend = "bottom")

a
b

```

# Baseline asymmetric: dominant due to ethnic ration

```{r asym_baseline, fig.cap="Fig. 8 orig", echo = FALSE}

asym_dom <- read.csv("df_asym_dom.csv",sep = ",")

excl <- names(asym_dom) %in% c("run", "density","step")
asym_dom <- asym_dom[!excl] %>% group_by(dominant,circle_orange,circle_blue,fraction_blue,secondary) %>% summarise_all(mean)

asym_dom <-  read.csv("df_asym_dom.csv",sep = ",") %>% select(-c(run,density,step)) %>% group_by(dominant,circle_orange,circle_blue,fraction_blue,secondary) %>% summarise_all(mean) %>% 
  gather(key = subgroup, value = Segra, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)  %>% mutate(
`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "Value" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"Ethnic"), 
 `ethnicity` = fct_relevel(fct_collapse(subgroup, "Majority" = c("de_sq_bl","dv_cl_bl","de_cl_bl","dv_sq_bl"), "Minority" = c("de_sq_or","de_cl_or","dv_sq_or","dv_cl_or")), "Majority"),
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "Liberals" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")), "Conservatives"), 
`type` = fct_relevel(fct_collapse(subgroup,"Conservative Majority" = c("de_sq_bl","dv_sq_bl"),
      "Liberals Majority" = c("de_cl_bl","dv_cl_bl"),
      "Conservative Minority" = c("de_sq_or","dv_sq_or"),
      "Liberals Minority" = c("de_cl_or","dv_cl_or")), "Conservative Majority")) %>%
  filter(fraction_blue == 50 | fraction_blue == 80)

asym_dom <- ggplot(data = asym_dom, aes(x = dominant, y = Segra, color = `segregation`, linetype = as.factor(fraction_blue))) + geom_point(size = 1.5) + geom_line() +
  facet_wrap(~ `type`, nrow = 1, scales= "free") +
  scale_color_manual(values = c("dark green","purple"),labels =  c("Ethnic","Value"),guide = guide_legend(title = "Segregation")) +
   scale_linetype_manual(values = c("solid","dashed"),labels =  c("50 %","80 %"),guide = guide_legend(title = "% Majority")) +
  ylab("Spatial Clustering") +
  xlab("ß dominant") +
    geom_hline(yintercept=1,color="red") +
  theme_bw() +
 theme(legend.position = "bottom")

ggsave("asym_dom.jpg",asym_dom, path = "images", width = 9, height = 4, dpi = 300)

```

# Comparing beta majority over beta minority liberal

```{r plot asym}

sens_vl <- read.csv("sens_vl.csv",sep =",") %>% select(-c(run)) %>%
  group_by(density,fraction_blue,circle_blue,circle_orange,eth_con_loc,eth_con_min,val_lib_loc,val_lib_min,eth_lib_loc,eth_lib_min,val_con_loc,val_con_min,ID) %>% summarise_all(mean) %>%  gather(key = subgroup, value = Segra, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)  %>% mutate(
`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "Value" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"Ethnic"), 
 `ethnicity` = fct_relevel(fct_collapse(subgroup, "Majority" = c("de_sq_bl","dv_cl_bl","de_cl_bl","dv_sq_bl"), "Minority" = c("de_sq_or","de_cl_or","dv_sq_or","dv_cl_or")), "Majority"),
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "Liberals" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")), "Conservatives"), 
`type` = fct_relevel(fct_collapse(subgroup,"Conservative Majority" = c("de_sq_bl","dv_sq_bl"),
      "Liberals Majority" = c("de_cl_bl","dv_cl_bl"),
      "Conservative Minority" = c("de_sq_or","dv_sq_or"),
      "Liberals Minority" = c("de_cl_or","dv_cl_or")), "Conservative Majority"))

asym.plot <- function(i){
  suppressWarnings(
ggplot(data = i, aes(x= if(i$ID == "mj"){i$val_lib_loc}else{i$val_lib_min}, y=Segra, color = segregation, linetype = as.factor(if (i$ID == "mj"){i$val_lib_min}else{i$val_lib_loc}))) + geom_point(size = 1.5) + geom_line() +
  facet_wrap(~ type, scales = "free") +
  scale_linetype_manual(values = c("solid","dotted","dashed"),labels = c("0","1","15"), guide = guide_legend(title = 
   if (i$ID == "mj"){ "ß liberal minority"}else{ "ß liberal majority"})) +
  scale_color_manual(values = c("dark green","purple"), labels =  c("ethnic","value"), guide = guide_legend(title = "Segregation")) +
  xlab(if (i$ID == "mj"){ "ß liberal majority"}else{ "ß liberal minority"}) +
  ylab("Spatial Clustering") +
  geom_hline(yintercept=1,color="red") +
  theme_bw() +
  theme(legend.position="right")
)
}

```

```{r asym liberals}

mnlib <- asym.plot( sens_vl %>%  filter( ID == "mn" & type == "Liberals Minority"))
mjlib <- asym.plot(sens_vl %>%  filter( ID == "mj" & type == "Liberals Majority"))

asym_lib <- suppressWarnings( ggpubr::ggarrange(mnlib,mjlib,ncol=1,nrow=2))

ggsave("asym_lib.jpg",asym_lib,path = "images")

```

```{r asym conservatives}

mncon <- asym.plot( sens_vl %>%  filter( type == "Conservative Majority" | type == "Conservative Minority") %>% filter(ID == "mn"))
mjcon <- asym.plot(sens_vl %>%  filter( type == "Conservative Majority" | type == "Conservative Minority") %>% filter(ID == "mj"))

asym_cons <- suppressWarnings( ggpubr::ggarrange(mncon,mjcon,ncol=1,nrow=2))

ggsave("asym_cons.jpg",asym_cons,path = "images")

```

```{r asym dislib}

dislib <- read.csv("dislib.csv",sep = ",") %>% select(-c(run)) %>%
  group_by(density,fraction_blue,circle_blue,circle_orange,eth_con_loc,eth_con_min,val_lib_loc,val_lib_min,eth_lib_loc,eth_lib_min,val_con_loc,val_con_min,ID) %>% summarise_all(mean) %>%  gather(key = subgroup, value = Segra, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)  %>% mutate(
`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "Value" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"Ethnic"), 
 `ethnicity` = fct_relevel(fct_collapse(subgroup, "Majority" = c("de_sq_bl","dv_cl_bl","de_cl_bl","dv_sq_bl"), "Minority" = c("de_sq_or","de_cl_or","dv_sq_or","dv_cl_or")), "Majority"),
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "Liberals" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")), "Conservatives"), 
`type` = fct_relevel(fct_collapse(subgroup,"Conservative Majority" = c("de_sq_bl","dv_sq_bl"),
      "Liberals Majority" = c("de_cl_bl","dv_cl_bl"),
      "Conservative Minority" = c("de_sq_or","dv_sq_or"),
      "Liberals Minority" = c("de_cl_or","dv_cl_or")), "Conservative Majority"))

ethnicity.labs <- c("50% Majority", "60% Majority","70% Majority","80% Majority","90% Majority")
names(ethnicity.labs) <- c("50","60","70","80","90")

st <- list(
dislib %>% filter(type == "Conservative Majority"),
dislib %>% filter(type == "Conservative Minority"),
dislib %>% filter(type == "Liberals Majority"),
dislib %>% filter(type == "Liberals Minority")

)

suppressWarnings(
for (df in st) {

pl <- ggplot(data = df, aes(x= circle_orange, y=Segra, color = segregation, linetype = as.factor(ID))) + geom_point(size=1.5) + geom_line() +
  facet_wrap(~ fraction_blue, nrow=1, labeller = labeller(fraction_blue = ethnicity.labs)) +
scale_linetype_manual(values = c(bsl = "solid", mj1 = "dashed", mn1 = "dotted"),labels = c("Basine (ß = 15)","ß liberal majority = 1","ß liberal minority = 1"), guide = guide_legend(title = "Dominant \nPreference:")) +
  scale_color_manual(values = c("dark green","purple"), labels =  c("ethnic","value"), guide = guide_legend(title = "Segregation")) +
  xlab("% Liberals Minority") +
  ylab("Spatial Clustering") +
   scale_x_discrete(limits = seq(10,90 ,by = 10)) +
  geom_hline(yintercept=1,color="red") +
  ggtitle(df$type) +
  theme_bw() +
  theme(panel.spacing = unit(0.7, "lines"),legend.position="bottom",legend.text=element_text(size=12))

nmn <- c(if(df$type == "Conservative Majority"){"dislib_cn_mj"}, if(df$type == "Conservative Minority"){"dislib_cn_mn"},if(df$type == "Liberals Majority"){"dislib_lb_mj"},if(df$type == "Liberals Minority"){"dislib_lb_mn"})

ggsave(paste0(nmn,".jpg"), pl,path = "images", width = 10, height = 6, dpi = 300)

}

)

```

# Increase ethnic liberal

```{r ethnicliberal, echo = FALSE}

etlb <- read.csv("etlb.csv",sep=",") %>% select(-c(run)) %>%
  group_by(density,fraction_blue,circle_blue,circle_orange,et_mn_on,et_mj_on,dominant,ID) %>% summarise_all(mean) %>%  gather(key = subgroup, value = Segra, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)  %>% mutate(
`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "Value" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"Ethnic"), 
 `ethnicity` = fct_relevel(fct_collapse(subgroup, "Majority" = c("de_sq_bl","dv_cl_bl","de_cl_bl","dv_sq_bl"), "Minority" = c("de_sq_or","de_cl_or","dv_sq_or","dv_cl_or")), "Majority"),
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "Liberals" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")), "Conservatives"), 
`type` = fct_relevel(fct_collapse(subgroup,"Conservative Majority" = c("de_sq_bl","dv_sq_bl"),
      "Liberals Majority" = c("de_cl_bl","dv_cl_bl"),
      "Conservative Minority" = c("de_sq_or","dv_sq_or"),
      "Liberals Minority" = c("de_cl_or","dv_cl_or")), "Conservative Majority")) %>% filter(fraction_blue == 50| fraction_blue == 80)

cl <- list(
as.data.frame(cbind(subset(etlb, type == "Conservative Majority" | type == "Conservative Minority"), s = "conservatives")),
as.data.frame(cbind(subset(etlb, type == "Liberals Majority" | type == "Liberals Minority"), s = "liberals"))
)

suppressWarnings(

for (i in cl) {

p <- ggplot(i, aes(dominant,y=Segra,color =`segregation`,linetype = as.factor(ID) )) +
    geom_point(size = 1.5, shape = if(i$s == "conservatives"){"square"}else{"circle"}) +
    geom_line() +
    facet_wrap(type ~ fraction_blue,  scales = "free", labeller = labeller(fraction_blue = c("50" = "50% Majority", "80" = "80% Majority"))) + 
    scale_color_manual(values = c("darkgreen","purple"), labels = c("Ethnic","Value"), guide_legend(title = "Segregation:")) +
        scale_linetype_manual(values = c("solid","longdash","dotted"), labels = c("Baseline","Lib Maj","Lib Min"), guide_legend(title = "Beta Ethnic:")) +
    ylab("Spatial Clustering") +
    xlab("ß dominant") +
    theme_bw() +
    theme(legend.position = "bottom")

  plot(p)

  nm <- paste0(as.name(i$s),".jpg")
  ggsave(nm,p,path = "images")

}

)

```

# Increase ethnic: distribution liberals

```{r dislib_et}

dislib_et <- read.csv("dislib_et.csv",sep=",") %>% select(-c(run)) %>%
  group_by(density,fraction_blue,circle_blue,circle_orange,et_mn_on,et_mj_on,dominant,ID) %>% summarise_all(mean) %>%  gather(key = subgroup, value = Segra, de_sq_bl,dv_sq_bl,de_cl_bl,dv_cl_bl,de_sq_or,dv_sq_or,de_cl_or,dv_cl_or)  %>% mutate(
`segregation` = fct_relevel(fct_collapse(subgroup,"Ethnic" = c("de_sq_bl","de_cl_bl","de_sq_or","de_cl_or"), "Value" = c("dv_sq_bl","dv_cl_bl","dv_sq_or","dv_cl_or")),"Ethnic"), 
 `ethnicity` = fct_relevel(fct_collapse(subgroup, "Majority" = c("de_sq_bl","dv_cl_bl","de_cl_bl","dv_sq_bl"), "Minority" = c("de_sq_or","de_cl_or","dv_sq_or","dv_cl_or")), "Majority"),
`orientation` = fct_relevel(fct_collapse(subgroup,"Conservatives" = c("de_sq_bl","dv_sq_bl","de_sq_or","dv_sq_or"), "Liberals" = c("de_cl_bl","dv_cl_bl","de_cl_or","dv_cl_or")), "Conservatives"), 
`type` = fct_relevel(fct_collapse(subgroup,"Conservative Majority" = c("de_sq_bl","dv_sq_bl"),
      "Liberals Majority" = c("de_cl_bl","dv_cl_bl"),
      "Conservative Minority" = c("de_sq_or","dv_sq_or"),
      "Liberals Minority" = c("de_cl_or","dv_cl_or")), "Conservative Majority"))

ethnicity.labs <- c("50% Majority", "60% Majority","70% Majority","80% Majority","90% Majority")
names(ethnicity.labs) <- c("50","60","70","80","90")

st <- list(
dislib_et %>% filter(type == "Conservative Majority"),
dislib_et %>% filter(type == "Conservative Minority"),
dislib_et %>% filter(type == "Liberals Majority"),
dislib_et %>% filter(type == "Liberals Minority")

)

suppressWarnings(
for (df in st) {

pl <- ggplot(data = df, aes(x= circle_orange, y=Segra, color = segregation, linetype = as.factor(ID))) + geom_point(size=1.5) + geom_line() +
  facet_wrap(~ fraction_blue, nrow=1, labeller = labeller(fraction_blue = ethnicity.labs)) +
scale_linetype_manual(values = c(bsl = "solid", etmj = "dashed", etmn = "dotted"),labels = c("Basine (ß = 0)","ß liberal majority","ß liberal minority"), guide = guide_legend(title = "ß ethnic")) +
  scale_color_manual(values = c("dark green","purple"), labels =  c("ethnic","value"), guide = guide_legend(title = "Segregation")) +
  xlab("% Liberals Minority") +
  ylab("Spatial Clustering") +
   scale_x_discrete(limits = seq(10,90 ,by = 10)) +
  geom_hline(yintercept=1,color="red") +
  ggtitle(df$type) +
  theme_bw() +
  theme(panel.spacing = unit(0.7, "lines"),legend.position="bottom",legend.text=element_text(size=12))

nmn <- c(if(df$type == "Conservative Majority"){"dislib_cn_mj_et"}, if(df$type == "Conservative Minority"){"dislib_cn_mn_et"},if(df$type == "Liberals Majority"){"dislib_lb_mj_et"},if(df$type == "Liberals Minority"){"dislib_lb_mn_et"})

ggsave(paste0(nmn,".jpg"), pl,path = "images", width = 10, height = 6, dpi = 300)

}

)

```